### 2021-04-05 theory/figures/fig1.tex

\documentclass{standalone}

\usepackage{tikz,pgf} %and any other packages or tikzlibraries your picture needs

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%COPY HERE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300

%Straight Lines [id:da9655093306166109]
\draw    (151.25,213.56) -- (420.92,213.12) ;
%Straight Lines [id:da6780197443194285]
\draw    (178.99,167.4) -- (393.4,167.4) ;
%Straight Lines [id:da9790479690343763]
\draw    (179.39,167) -- (151.25,213.56) ;
%Straight Lines [id:da9225563682209392]
\draw    (393.8,167) -- (420.92,213.12) ;
%Straight Lines [id:da09563387136808865]
\draw    (286.6,167) -- (286.08,213.34) ;
%Straight Lines [id:da3116522709021452]
\draw    (273.92,166.96) -- (270.68,213.56) ;
%Straight Lines [id:da7104288201908657]
\draw    (259.41,166.96) -- (253.18,213.3) ;
%Straight Lines [id:da742779405875688]
\draw    (245.3,166.91) -- (237.52,213.3) ;
%Straight Lines [id:da18844905298889292]
\draw    (232.1,167.18) -- (220.33,213.82) ;
%Straight Lines [id:da35804353174173054]
\draw    (217.96,167.2) -- (201.3,213.56) ;
%Straight Lines [id:da7688901048445682]
\draw    (205.7,167.18) -- (185.02,213.82) ;
%Straight Lines [id:da3470294213255083]
\draw    (192.98,167.2) -- (169.98,213.82) ;
%Straight Lines [id:da35862455135349536]
\draw    (298.62,166.91) -- (302.61,213.56) ;
%Straight Lines [id:da29999796293225955]
\draw    (313.18,167.08) -- (321.21,213.92) ;
%Straight Lines [id:da6206528590435849]
\draw    (327.48,166.91) -- (339.76,213.3) ;
%Straight Lines [id:da9408253594123648]
\draw    (340.68,167.18) -- (357.57,213.82) ;
%Straight Lines [id:da917785587155304]
\draw    (354.8,167.18) -- (374.76,213.56) ;
%Straight Lines [id:da3801603892210079]
\draw    (366.16,167.08) -- (392.26,213.56) ;
%Straight Lines [id:da580723776969265]
\draw    (380.59,167.18) -- (405.77,213.3) ;
%Straight Lines [id:da5336171972507551]
\draw    (155.55,205.39) -- (416.51,204.6) ;
%Straight Lines [id:da956646312420568]
\draw    (160.77,197.48) -- (411.6,196.96) ;
%Straight Lines [id:da2531401201560137]
\draw    (165.32,190.28) -- (407.36,190.06) ;
%Straight Lines [id:da31575285130833475]
\draw    (169.06,183.52) -- (403.31,183.25) ;
%Straight Lines [id:da6000067635957629]
\draw    (172.65,177.85) -- (399.54,177.33) ;
%Straight Lines [id:da36539132370134153]
\draw    (175.41,172.58) -- (396.16,171.79) ;
%Shape: Ellipse [id:dp12271593300544881]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (172.44,205.39) .. controls (172.44,204.59) and (173.19,203.94) .. (174.12,203.94) .. controls (175.06,203.94) and (175.81,204.59) .. (175.81,205.39) .. controls (175.81,206.19) and (175.06,206.84) .. (174.12,206.84) .. controls (173.19,206.84) and (172.44,206.19) .. (172.44,205.39) -- cycle ;
%Shape: Ellipse [id:dp17187409527803665]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (176.2,197.36) .. controls (176.2,196.56) and (176.96,195.91) .. (177.89,195.91) .. controls (178.82,195.91) and (179.58,196.56) .. (179.58,197.36) .. controls (179.58,198.16) and (178.82,198.81) .. (177.89,198.81) .. controls (176.96,198.81) and (176.2,198.16) .. (176.2,197.36) -- cycle ;
%Shape: Ellipse [id:dp3897463489917039]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (179.83,189.94) .. controls (179.83,189.14) and (180.59,188.49) .. (181.52,188.49) .. controls (182.45,188.49) and (183.21,189.14) .. (183.21,189.94) .. controls (183.21,190.74) and (182.45,191.39) .. (181.52,191.39) .. controls (180.59,191.39) and (179.83,190.74) .. (179.83,189.94) -- cycle ;
%Shape: Ellipse [id:dp6651227824721306]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (182.9,183.59) .. controls (182.9,182.79) and (183.66,182.14) .. (184.59,182.14) .. controls (185.52,182.14) and (186.28,182.79) .. (186.28,183.59) .. controls (186.28,184.39) and (185.52,185.04) .. (184.59,185.04) .. controls (183.66,185.04) and (182.9,184.39) .. (182.9,183.59) -- cycle ;
%Shape: Ellipse [id:dp16931876417539748]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (186.11,177.24) .. controls (186.11,176.44) and (186.87,175.79) .. (187.8,175.79) .. controls (188.73,175.79) and (189.49,176.44) .. (189.49,177.24) .. controls (189.49,178.04) and (188.73,178.69) .. (187.8,178.69) .. controls (186.87,178.69) and (186.11,178.04) .. (186.11,177.24) -- cycle ;
%Shape: Ellipse [id:dp011630715757735333]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (189.04,171.85) .. controls (189.04,171.05) and (189.8,170.4) .. (190.73,170.4) .. controls (191.66,170.4) and (192.42,171.05) .. (192.42,171.85) .. controls (192.42,172.65) and (191.66,173.3) .. (190.73,173.3) .. controls (189.8,173.3) and (189.04,172.65) .. (189.04,171.85) -- cycle ;
%Shape: Ellipse [id:dp6352732828874232]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (186.95,205.15) .. controls (186.95,204.35) and (187.71,203.7) .. (188.64,203.7) .. controls (189.57,203.7) and (190.33,204.35) .. (190.33,205.15) .. controls (190.33,205.95) and (189.57,206.6) .. (188.64,206.6) .. controls (187.71,206.6) and (186.95,205.95) .. (186.95,205.15) -- cycle ;
%Shape: Ellipse [id:dp5586918813010036]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (190.44,197.48) .. controls (190.44,196.68) and (191.19,196.03) .. (192.13,196.03) .. controls (193.06,196.03) and (193.82,196.68) .. (193.82,197.48) .. controls (193.82,198.28) and (193.06,198.93) .. (192.13,198.93) .. controls (191.19,198.93) and (190.44,198.28) .. (190.44,197.48) -- cycle ;
%Shape: Ellipse [id:dp9378136236695127]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (193.67,190.5) .. controls (193.67,189.7) and (194.43,189.05) .. (195.36,189.05) .. controls (196.29,189.05) and (197.05,189.7) .. (197.05,190.5) .. controls (197.05,191.3) and (196.29,191.95) .. (195.36,191.95) .. controls (194.43,191.95) and (193.67,191.3) .. (193.67,190.5) -- cycle ;
%Shape: Ellipse [id:dp7452150293769777]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (196.58,183.47) .. controls (196.58,182.67) and (197.33,182.02) .. (198.27,182.02) .. controls (199.2,182.02) and (199.96,182.67) .. (199.96,183.47) .. controls (199.96,184.27) and (199.2,184.92) .. (198.27,184.92) .. controls (197.33,184.92) and (196.58,184.27) .. (196.58,183.47) -- cycle ;
%Shape: Ellipse [id:dp2595074236120878]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (199.51,177.48) .. controls (199.51,176.68) and (200.27,176.03) .. (201.2,176.03) .. controls (202.13,176.03) and (202.89,176.68) .. (202.89,177.48) .. controls (202.89,178.28) and (202.13,178.93) .. (201.2,178.93) .. controls (200.27,178.93) and (199.51,178.28) .. (199.51,177.48) -- cycle ;
%Shape: Ellipse [id:dp8182709430461124]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (201.6,171.85) .. controls (201.6,171.05) and (202.36,170.4) .. (203.29,170.4) .. controls (204.22,170.4) and (204.98,171.05) .. (204.98,171.85) .. controls (204.98,172.65) and (204.22,173.3) .. (203.29,173.3) .. controls (202.36,173.3) and (201.6,172.65) .. (201.6,171.85) -- cycle ;
%Shape: Ellipse [id:dp8721800205864367]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (202.72,205.27) .. controls (202.72,204.47) and (203.47,203.82) .. (204.41,203.82) .. controls (205.34,203.82) and (206.1,204.47) .. (206.1,205.27) .. controls (206.1,206.07) and (205.34,206.72) .. (204.41,206.72) .. controls (203.47,206.72) and (202.72,206.07) .. (202.72,205.27) -- cycle ;
%Shape: Ellipse [id:dp7821722952675314]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (205.51,197.36) .. controls (205.51,196.56) and (206.27,195.91) .. (207.2,195.91) .. controls (208.13,195.91) and (208.89,196.56) .. (208.89,197.36) .. controls (208.89,198.16) and (208.13,198.81) .. (207.2,198.81) .. controls (206.27,198.81) and (205.51,198.16) .. (205.51,197.36) -- cycle ;
%Shape: Ellipse [id:dp6361699950198552]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (207.95,190.5) .. controls (207.95,189.7) and (208.7,189.05) .. (209.64,189.05) .. controls (210.57,189.05) and (211.32,189.7) .. (211.32,190.5) .. controls (211.32,191.3) and (210.57,191.95) .. (209.64,191.95) .. controls (208.7,191.95) and (207.95,191.3) .. (207.95,190.5) -- cycle ;
%Shape: Ellipse [id:dp07429235889252994]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (210.53,183.35) .. controls (210.53,182.55) and (211.29,181.9) .. (212.22,181.9) .. controls (213.15,181.9) and (213.91,182.55) .. (213.91,183.35) .. controls (213.91,184.15) and (213.15,184.8) .. (212.22,184.8) .. controls (211.29,184.8) and (210.53,184.15) .. (210.53,183.35) -- cycle ;
%Shape: Ellipse [id:dp46939239369131425]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (212.63,177.12) .. controls (212.63,176.32) and (213.38,175.67) .. (214.32,175.67) .. controls (215.25,175.67) and (216,176.32) .. (216,177.12) .. controls (216,177.92) and (215.25,178.57) .. (214.32,178.57) .. controls (213.38,178.57) and (212.63,177.92) .. (212.63,177.12) -- cycle ;
%Shape: Ellipse [id:dp47839368197126064]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (214.58,171.73) .. controls (214.58,170.93) and (215.34,170.28) .. (216.27,170.28) .. controls (217.2,170.28) and (217.96,170.93) .. (217.96,171.73) .. controls (217.96,172.53) and (217.2,173.18) .. (216.27,173.18) .. controls (215.34,173.18) and (214.58,172.53) .. (214.58,171.73) -- cycle ;
%Shape: Ellipse [id:dp6735819197452304]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (220.72,205.39) .. controls (220.72,204.59) and (221.48,203.94) .. (222.41,203.94) .. controls (223.34,203.94) and (224.1,204.59) .. (224.1,205.39) .. controls (224.1,206.19) and (223.34,206.84) .. (222.41,206.84) .. controls (221.48,206.84) and (220.72,206.19) .. (220.72,205.39) -- cycle ;
%Shape: Ellipse [id:dp40045376893902485]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (222.81,197.48) .. controls (222.81,196.68) and (223.57,196.03) .. (224.5,196.03) .. controls (225.44,196.03) and (226.19,196.68) .. (226.19,197.48) .. controls (226.19,198.28) and (225.44,198.93) .. (224.5,198.93) .. controls (223.57,198.93) and (222.81,198.28) .. (222.81,197.48) -- cycle ;
%Shape: Ellipse [id:dp002588008650496576]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (224.53,190.5) .. controls (224.53,189.7) and (225.28,189.05) .. (226.21,189.05) .. controls (227.15,189.05) and (227.9,189.7) .. (227.9,190.5) .. controls (227.9,191.3) and (227.15,191.95) .. (226.21,191.95) .. controls (225.28,191.95) and (224.53,191.3) .. (224.53,190.5) -- cycle ;
%Shape: Ellipse [id:dp7269110513068167]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (226.16,183.59) .. controls (226.16,182.79) and (226.92,182.14) .. (227.85,182.14) .. controls (228.78,182.14) and (229.54,182.79) .. (229.54,183.59) .. controls (229.54,184.39) and (228.78,185.04) .. (227.85,185.04) .. controls (226.92,185.04) and (226.16,184.39) .. (226.16,183.59) -- cycle ;
%Shape: Ellipse [id:dp6228752312232548]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (227.7,177.36) .. controls (227.7,176.56) and (228.45,175.91) .. (229.39,175.91) .. controls (230.32,175.91) and (231.08,176.56) .. (231.08,177.36) .. controls (231.08,178.16) and (230.32,178.81) .. (229.39,178.81) .. controls (228.45,178.81) and (227.7,178.16) .. (227.7,177.36) -- cycle ;
%Shape: Ellipse [id:dp5712591071424813]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (229.09,171.97) .. controls (229.09,171.17) and (229.85,170.52) .. (230.78,170.52) .. controls (231.72,170.52) and (232.47,171.17) .. (232.47,171.97) .. controls (232.47,172.77) and (231.72,173.42) .. (230.78,173.42) .. controls (229.85,173.42) and (229.09,172.77) .. (229.09,171.97) -- cycle ;
%Shape: Ellipse [id:dp37415663269795374]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (242.77,171.85) .. controls (242.77,171.05) and (243.53,170.4) .. (244.46,170.4) .. controls (245.39,170.4) and (246.15,171.05) .. (246.15,171.85) .. controls (246.15,172.65) and (245.39,173.3) .. (244.46,173.3) .. controls (243.53,173.3) and (242.77,172.65) .. (242.77,171.85) -- cycle ;
%Shape: Ellipse [id:dp5844680482576305]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (241.79,177.48) .. controls (241.79,176.68) and (242.55,176.03) .. (243.48,176.03) .. controls (244.41,176.03) and (245.17,176.68) .. (245.17,177.48) .. controls (245.17,178.28) and (244.41,178.93) .. (243.48,178.93) .. controls (242.55,178.93) and (241.79,178.28) .. (241.79,177.48) -- cycle ;
%Shape: Ellipse [id:dp27800840767464763]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (240.68,183.23) .. controls (240.68,182.43) and (241.43,181.78) .. (242.37,181.78) .. controls (243.3,181.78) and (244.05,182.43) .. (244.05,183.23) .. controls (244.05,184.03) and (243.3,184.68) .. (242.37,184.68) .. controls (241.43,184.68) and (240.68,184.03) .. (240.68,183.23) -- cycle ;
%Shape: Ellipse [id:dp6309981882755835]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (239.72,190.1) .. controls (239.72,189.3) and (240.48,188.65) .. (241.41,188.65) .. controls (242.34,188.65) and (243.1,189.3) .. (243.1,190.1) .. controls (243.1,190.9) and (242.34,191.55) .. (241.41,191.55) .. controls (240.48,191.55) and (239.72,190.9) .. (239.72,190.1) -- cycle ;
%Shape: Ellipse [id:dp6064458479575909]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (238.44,197.36) .. controls (238.44,196.56) and (239.2,195.91) .. (240.13,195.91) .. controls (241.07,195.91) and (241.82,196.56) .. (241.82,197.36) .. controls (241.82,198.16) and (241.07,198.81) .. (240.13,198.81) .. controls (239.2,198.81) and (238.44,198.16) .. (238.44,197.36) -- cycle ;
%Shape: Ellipse [id:dp8566892005164177]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (237.19,205.03) .. controls (237.19,204.23) and (237.94,203.58) .. (238.88,203.58) .. controls (239.81,203.58) and (240.57,204.23) .. (240.57,205.03) .. controls (240.57,205.83) and (239.81,206.48) .. (238.88,206.48) .. controls (237.94,206.48) and (237.19,205.83) .. (237.19,205.03) -- cycle ;
%Shape: Ellipse [id:dp21811991550458654]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (257.01,171.85) .. controls (257.01,171.05) and (257.76,170.4) .. (258.69,170.4) .. controls (259.63,170.4) and (260.38,171.05) .. (260.38,171.85) .. controls (260.38,172.65) and (259.63,173.3) .. (258.69,173.3) .. controls (257.76,173.3) and (257.01,172.65) .. (257.01,171.85) -- cycle ;
%Shape: Ellipse [id:dp07465207554443909]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (256.17,177.36) .. controls (256.17,176.56) and (256.92,175.91) .. (257.86,175.91) .. controls (258.79,175.91) and (259.54,176.56) .. (259.54,177.36) .. controls (259.54,178.16) and (258.79,178.81) .. (257.86,178.81) .. controls (256.92,178.81) and (256.17,178.16) .. (256.17,177.36) -- cycle ;
%Shape: Ellipse [id:dp3520799274177444]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (255.33,183.47) .. controls (255.33,182.67) and (256.09,182.02) .. (257.02,182.02) .. controls (257.95,182.02) and (258.71,182.67) .. (258.71,183.47) .. controls (258.71,184.27) and (257.95,184.92) .. (257.02,184.92) .. controls (256.09,184.92) and (255.33,184.27) .. (255.33,183.47) -- cycle ;
%Shape: Ellipse [id:dp8618279200116534]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (254.56,190.37) .. controls (254.56,189.57) and (255.32,188.92) .. (256.25,188.92) .. controls (257.18,188.92) and (257.94,189.57) .. (257.94,190.37) .. controls (257.94,191.17) and (257.18,191.82) .. (256.25,191.82) .. controls (255.32,191.82) and (254.56,191.17) .. (254.56,190.37) -- cycle ;
%Shape: Ellipse [id:dp4846168731141014]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (253.52,197.24) .. controls (253.52,196.44) and (254.27,195.79) .. (255.2,195.79) .. controls (256.14,195.79) and (256.89,196.44) .. (256.89,197.24) .. controls (256.89,198.04) and (256.14,198.69) .. (255.2,198.69) .. controls (254.27,198.69) and (253.52,198.04) .. (253.52,197.24) -- cycle ;
%Shape: Ellipse [id:dp3684815163642654]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (252.4,205.15) .. controls (252.4,204.35) and (253.16,203.7) .. (254.09,203.7) .. controls (255.02,203.7) and (255.78,204.35) .. (255.78,205.15) .. controls (255.78,205.95) and (255.02,206.6) .. (254.09,206.6) .. controls (253.16,206.6) and (252.4,205.95) .. (252.4,205.15) -- cycle ;
%Shape: Ellipse [id:dp1437934667713081]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (271.94,171.85) .. controls (271.94,171.05) and (272.69,170.4) .. (273.63,170.4) .. controls (274.56,170.4) and (275.31,171.05) .. (275.31,171.85) .. controls (275.31,172.65) and (274.56,173.3) .. (273.63,173.3) .. controls (272.69,173.3) and (271.94,172.65) .. (271.94,171.85) -- cycle ;
%Shape: Ellipse [id:dp886301588499915]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (271.52,177.24) .. controls (271.52,176.44) and (272.27,175.79) .. (273.21,175.79) .. controls (274.14,175.79) and (274.9,176.44) .. (274.9,177.24) .. controls (274.9,178.04) and (274.14,178.69) .. (273.21,178.69) .. controls (272.27,178.69) and (271.52,178.04) .. (271.52,177.24) -- cycle ;
%Shape: Ellipse [id:dp10861025652876077]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (271.24,183.23) .. controls (271.24,182.43) and (272,181.78) .. (272.93,181.78) .. controls (273.86,181.78) and (274.62,182.43) .. (274.62,183.23) .. controls (274.62,184.03) and (273.86,184.68) .. (272.93,184.68) .. controls (272,184.68) and (271.24,184.03) .. (271.24,183.23) -- cycle ;
%Shape: Ellipse [id:dp8227079118183289]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (270.68,190.5) .. controls (270.68,189.7) and (271.44,189.05) .. (272.37,189.05) .. controls (273.3,189.05) and (274.06,189.7) .. (274.06,190.5) .. controls (274.06,191.3) and (273.3,191.95) .. (272.37,191.95) .. controls (271.44,191.95) and (270.68,191.3) .. (270.68,190.5) -- cycle ;
%Shape: Ellipse [id:dp31754679610416936]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (270.12,197.12) .. controls (270.12,196.32) and (270.88,195.67) .. (271.81,195.67) .. controls (272.74,195.67) and (273.5,196.32) .. (273.5,197.12) .. controls (273.5,197.92) and (272.74,198.57) .. (271.81,198.57) .. controls (270.88,198.57) and (270.12,197.92) .. (270.12,197.12) -- cycle ;
%Shape: Ellipse [id:dp228275147103431]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (269.7,205.03) .. controls (269.7,204.23) and (270.46,203.58) .. (271.39,203.58) .. controls (272.33,203.58) and (273.08,204.23) .. (273.08,205.03) .. controls (273.08,205.83) and (272.33,206.48) .. (271.39,206.48) .. controls (270.46,206.48) and (269.7,205.83) .. (269.7,205.03) -- cycle ;
%Shape: Ellipse [id:dp5867689834208971]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (284.5,171.79) .. controls (284.5,170.99) and (285.25,170.34) .. (286.19,170.34) .. controls (287.12,170.34) and (287.87,170.99) .. (287.87,171.79) .. controls (287.87,172.59) and (287.12,173.24) .. (286.19,173.24) .. controls (285.25,173.24) and (284.5,172.59) .. (284.5,171.79) -- cycle ;
%Shape: Ellipse [id:dp18205732240558015]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (284.8,177.19) .. controls (284.8,176.39) and (285.56,175.74) .. (286.49,175.74) .. controls (287.43,175.74) and (288.18,176.39) .. (288.18,177.19) .. controls (288.18,177.99) and (287.43,178.64) .. (286.49,178.64) .. controls (285.56,178.64) and (284.8,177.99) .. (284.8,177.19) -- cycle ;
%Shape: Ellipse [id:dp6234929744101898]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (284.5,183.38) .. controls (284.5,182.58) and (285.25,181.93) .. (286.19,181.93) .. controls (287.12,181.93) and (287.87,182.58) .. (287.87,183.38) .. controls (287.87,184.18) and (287.12,184.83) .. (286.19,184.83) .. controls (285.25,184.83) and (284.5,184.18) .. (284.5,183.38) -- cycle ;
%Shape: Ellipse [id:dp8290479654576126]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (284.65,190.17) .. controls (284.65,189.37) and (285.41,188.72) .. (286.34,188.72) .. controls (287.27,188.72) and (288.03,189.37) .. (288.03,190.17) .. controls (288.03,190.97) and (287.27,191.62) .. (286.34,191.62) .. controls (285.41,191.62) and (284.65,190.97) .. (284.65,190.17) -- cycle ;
%Shape: Ellipse [id:dp8989481541795088]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (284.5,197.22) .. controls (284.5,196.42) and (285.25,195.77) .. (286.19,195.77) .. controls (287.12,195.77) and (287.87,196.42) .. (287.87,197.22) .. controls (287.87,198.02) and (287.12,198.67) .. (286.19,198.67) .. controls (285.25,198.67) and (284.5,198.02) .. (284.5,197.22) -- cycle ;
%Shape: Ellipse [id:dp3354120737649453]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (284.34,204.99) .. controls (284.34,204.19) and (285.1,203.54) .. (286.03,203.54) .. controls (286.96,203.54) and (287.72,204.19) .. (287.72,204.99) .. controls (287.72,205.79) and (286.96,206.44) .. (286.03,206.44) .. controls (285.1,206.44) and (284.34,205.79) .. (284.34,204.99) -- cycle ;
%Shape: Ellipse [id:dp3606226368237744]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (297.2,171.73) .. controls (297.2,170.93) and (297.95,170.28) .. (298.89,170.28) .. controls (299.82,170.28) and (300.57,170.93) .. (300.57,171.73) .. controls (300.57,172.53) and (299.82,173.18) .. (298.89,173.18) .. controls (297.95,173.18) and (297.2,172.53) .. (297.2,171.73) -- cycle ;
%Shape: Ellipse [id:dp048913958034914495]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (297.89,177.24) .. controls (297.89,176.44) and (298.65,175.79) .. (299.58,175.79) .. controls (300.52,175.79) and (301.27,176.44) .. (301.27,177.24) .. controls (301.27,178.04) and (300.52,178.69) .. (299.58,178.69) .. controls (298.65,178.69) and (297.89,178.04) .. (297.89,177.24) -- cycle ;
%Shape: Ellipse [id:dp19686667746567443]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (298.17,183.47) .. controls (298.17,182.67) and (298.93,182.02) .. (299.86,182.02) .. controls (300.79,182.02) and (301.55,182.67) .. (301.55,183.47) .. controls (301.55,184.27) and (300.79,184.92) .. (299.86,184.92) .. controls (298.93,184.92) and (298.17,184.27) .. (298.17,183.47) -- cycle ;
%Shape: Ellipse [id:dp4115460327510696]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (298.93,190.24) .. controls (298.93,189.44) and (299.68,188.79) .. (300.62,188.79) .. controls (301.55,188.79) and (302.3,189.44) .. (302.3,190.24) .. controls (302.3,191.04) and (301.55,191.69) .. (300.62,191.69) .. controls (299.68,191.69) and (298.93,191.04) .. (298.93,190.24) -- cycle ;
%Shape: Ellipse [id:dp19847266839128408]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (299.29,197.12) .. controls (299.29,196.32) and (300.05,195.67) .. (300.98,195.67) .. controls (301.91,195.67) and (302.67,196.32) .. (302.67,197.12) .. controls (302.67,197.92) and (301.91,198.57) .. (300.98,198.57) .. controls (300.05,198.57) and (299.29,197.92) .. (299.29,197.12) -- cycle ;
%Shape: Ellipse [id:dp48996962931873567]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (300.13,204.91) .. controls (300.13,204.11) and (300.88,203.46) .. (301.82,203.46) .. controls (302.75,203.46) and (303.5,204.11) .. (303.5,204.91) .. controls (303.5,205.71) and (302.75,206.36) .. (301.82,206.36) .. controls (300.88,206.36) and (300.13,205.71) .. (300.13,204.91) -- cycle ;
%Shape: Ellipse [id:dp8365501880357185]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (311.87,172.13) .. controls (311.87,171.33) and (312.62,170.68) .. (313.56,170.68) .. controls (314.49,170.68) and (315.25,171.33) .. (315.25,172.13) .. controls (315.25,172.93) and (314.49,173.58) .. (313.56,173.58) .. controls (312.62,173.58) and (311.87,172.93) .. (311.87,172.13) -- cycle ;
%Shape: Ellipse [id:dp3988926174843601]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (313.51,177.36) .. controls (313.51,176.56) and (314.26,175.91) .. (315.19,175.91) .. controls (316.13,175.91) and (316.88,176.56) .. (316.88,177.36) .. controls (316.88,178.16) and (316.13,178.81) .. (315.19,178.81) .. controls (314.26,178.81) and (313.51,178.16) .. (313.51,177.36) -- cycle ;
%Shape: Ellipse [id:dp04101225467897396]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (314.36,183.23) .. controls (314.36,182.43) and (315.12,181.78) .. (316.05,181.78) .. controls (316.98,181.78) and (317.74,182.43) .. (317.74,183.23) .. controls (317.74,184.03) and (316.98,184.68) .. (316.05,184.68) .. controls (315.12,184.68) and (314.36,184.03) .. (314.36,183.23) -- cycle ;
%Shape: Ellipse [id:dp5863915656225569]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (315.51,190.5) .. controls (315.51,189.7) and (316.26,189.05) .. (317.19,189.05) .. controls (318.13,189.05) and (318.88,189.7) .. (318.88,190.5) .. controls (318.88,191.3) and (318.13,191.95) .. (317.19,191.95) .. controls (316.26,191.95) and (315.51,191.3) .. (315.51,190.5) -- cycle ;
%Shape: Ellipse [id:dp17122444124667857]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (316.32,197.12) .. controls (316.32,196.32) and (317.07,195.67) .. (318,195.67) .. controls (318.94,195.67) and (319.69,196.32) .. (319.69,197.12) .. controls (319.69,197.92) and (318.94,198.57) .. (318,198.57) .. controls (317.07,198.57) and (316.32,197.92) .. (316.32,197.12) -- cycle ;
%Shape: Ellipse [id:dp595550673895106]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (317.99,205.03) .. controls (317.99,204.23) and (318.75,203.58) .. (319.68,203.58) .. controls (320.61,203.58) and (321.37,204.23) .. (321.37,205.03) .. controls (321.37,205.83) and (320.61,206.48) .. (319.68,206.48) .. controls (318.75,206.48) and (317.99,205.83) .. (317.99,205.03) -- cycle ;
%Shape: Ellipse [id:dp42805717802208254]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (326.92,171.49) .. controls (326.92,170.69) and (327.68,170.04) .. (328.61,170.04) .. controls (329.54,170.04) and (330.3,170.69) .. (330.3,171.49) .. controls (330.3,172.29) and (329.54,172.94) .. (328.61,172.94) .. controls (327.68,172.94) and (326.92,172.29) .. (326.92,171.49) -- cycle ;
%Shape: Ellipse [id:dp07574640518443809]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (328.6,177.24) .. controls (328.6,176.44) and (329.35,175.79) .. (330.28,175.79) .. controls (331.22,175.79) and (331.97,176.44) .. (331.97,177.24) .. controls (331.97,178.04) and (331.22,178.69) .. (330.28,178.69) .. controls (329.35,178.69) and (328.6,178.04) .. (328.6,177.24) -- cycle ;
%Shape: Ellipse [id:dp35961760720035785]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (329.99,183.35) .. controls (329.99,182.55) and (330.75,181.9) .. (331.68,181.9) .. controls (332.61,181.9) and (333.37,182.55) .. (333.37,183.35) .. controls (333.37,184.15) and (332.61,184.8) .. (331.68,184.8) .. controls (330.75,184.8) and (329.99,184.15) .. (329.99,183.35) -- cycle ;
%Shape: Ellipse [id:dp6949164860668007]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (331.93,190.1) .. controls (331.93,189.3) and (332.69,188.65) .. (333.62,188.65) .. controls (334.55,188.65) and (335.31,189.3) .. (335.31,190.1) .. controls (335.31,190.9) and (334.55,191.55) .. (333.62,191.55) .. controls (332.69,191.55) and (331.93,190.9) .. (331.93,190.1) -- cycle ;
%Shape: Ellipse [id:dp6551927674212119]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (333.76,197.24) .. controls (333.76,196.44) and (334.52,195.79) .. (335.45,195.79) .. controls (336.38,195.79) and (337.14,196.44) .. (337.14,197.24) .. controls (337.14,198.04) and (336.38,198.69) .. (335.45,198.69) .. controls (334.52,198.69) and (333.76,198.04) .. (333.76,197.24) -- cycle ;
%Shape: Ellipse [id:dp5972824761324165]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (335.71,205.03) .. controls (335.71,204.23) and (336.47,203.58) .. (337.4,203.58) .. controls (338.33,203.58) and (339.09,204.23) .. (339.09,205.03) .. controls (339.09,205.83) and (338.33,206.48) .. (337.4,206.48) .. controls (336.47,206.48) and (335.71,205.83) .. (335.71,205.03) -- cycle ;
%Shape: Ellipse [id:dp25895708663696104]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (340.46,171.49) .. controls (340.46,170.69) and (341.21,170.04) .. (342.15,170.04) .. controls (343.08,170.04) and (343.84,170.69) .. (343.84,171.49) .. controls (343.84,172.29) and (343.08,172.94) .. (342.15,172.94) .. controls (341.21,172.94) and (340.46,172.29) .. (340.46,171.49) -- cycle ;
%Shape: Ellipse [id:dp776678914241709]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (342.69,177) .. controls (342.69,176.2) and (343.45,175.55) .. (344.38,175.55) .. controls (345.31,175.55) and (346.07,176.2) .. (346.07,177) .. controls (346.07,177.8) and (345.31,178.45) .. (344.38,178.45) .. controls (343.45,178.45) and (342.69,177.8) .. (342.69,177) -- cycle ;
%Shape: Ellipse [id:dp7372524459660503]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (344.92,183.35) .. controls (344.92,182.55) and (345.68,181.9) .. (346.61,181.9) .. controls (347.54,181.9) and (348.3,182.55) .. (348.3,183.35) .. controls (348.3,184.15) and (347.54,184.8) .. (346.61,184.8) .. controls (345.68,184.8) and (344.92,184.15) .. (344.92,183.35) -- cycle ;
%Shape: Ellipse [id:dp24122491835471704]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (347.44,190.5) .. controls (347.44,189.7) and (348.19,189.05) .. (349.12,189.05) .. controls (350.06,189.05) and (350.81,189.7) .. (350.81,190.5) .. controls (350.81,191.3) and (350.06,191.95) .. (349.12,191.95) .. controls (348.19,191.95) and (347.44,191.3) .. (347.44,190.5) -- cycle ;
%Shape: Ellipse [id:dp7904451243351014]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (349.53,197) .. controls (349.53,196.2) and (350.28,195.55) .. (351.22,195.55) .. controls (352.15,195.55) and (352.91,196.2) .. (352.91,197) .. controls (352.91,197.8) and (352.15,198.45) .. (351.22,198.45) .. controls (350.28,198.45) and (349.53,197.8) .. (349.53,197) -- cycle ;
%Shape: Ellipse [id:dp4617144473175392]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (352.46,204.67) .. controls (352.46,203.87) and (353.22,203.22) .. (354.15,203.22) .. controls (355.08,203.22) and (355.84,203.87) .. (355.84,204.67) .. controls (355.84,205.47) and (355.08,206.12) .. (354.15,206.12) .. controls (353.22,206.12) and (352.46,205.47) .. (352.46,204.67) -- cycle ;
%Shape: Ellipse [id:dp626912606919074]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (354.97,171.49) .. controls (354.97,170.69) and (355.73,170.04) .. (356.66,170.04) .. controls (357.59,170.04) and (358.35,170.69) .. (358.35,171.49) .. controls (358.35,172.29) and (357.59,172.94) .. (356.66,172.94) .. controls (355.73,172.94) and (354.97,172.29) .. (354.97,171.49) -- cycle ;
%Shape: Ellipse [id:dp4306093121189267]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (357.48,177.12) .. controls (357.48,176.32) and (358.24,175.67) .. (359.17,175.67) .. controls (360.1,175.67) and (360.86,176.32) .. (360.86,177.12) .. controls (360.86,177.92) and (360.1,178.57) .. (359.17,178.57) .. controls (358.24,178.57) and (357.48,177.92) .. (357.48,177.12) -- cycle ;
%Shape: Ellipse [id:dp4919285418842567]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (359.86,183.23) .. controls (359.86,182.43) and (360.61,181.78) .. (361.54,181.78) .. controls (362.48,181.78) and (363.23,182.43) .. (363.23,183.23) .. controls (363.23,184.03) and (362.48,184.68) .. (361.54,184.68) .. controls (360.61,184.68) and (359.86,184.03) .. (359.86,183.23) -- cycle ;
%Shape: Ellipse [id:dp5744075183160424]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (362.65,189.82) .. controls (362.65,189.02) and (363.4,188.37) .. (364.34,188.37) .. controls (365.27,188.37) and (366.02,189.02) .. (366.02,189.82) .. controls (366.02,190.62) and (365.27,191.27) .. (364.34,191.27) .. controls (363.4,191.27) and (362.65,190.62) .. (362.65,189.82) -- cycle ;
%Shape: Ellipse [id:dp046203720018062944]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (366,197) .. controls (366,196.2) and (366.75,195.55) .. (367.68,195.55) .. controls (368.62,195.55) and (369.37,196.2) .. (369.37,197) .. controls (369.37,197.8) and (368.62,198.45) .. (367.68,198.45) .. controls (366.75,198.45) and (366,197.8) .. (366,197) -- cycle ;
%Shape: Ellipse [id:dp49972425736995807]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (369.21,204.67) .. controls (369.21,203.87) and (369.96,203.22) .. (370.89,203.22) .. controls (371.83,203.22) and (372.58,203.87) .. (372.58,204.67) .. controls (372.58,205.47) and (371.83,206.12) .. (370.89,206.12) .. controls (369.96,206.12) and (369.21,205.47) .. (369.21,204.67) -- cycle ;
%Shape: Ellipse [id:dp5024602360079904]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (367.39,171.49) .. controls (367.39,170.69) and (368.15,170.04) .. (369.08,170.04) .. controls (370.01,170.04) and (370.77,170.69) .. (370.77,171.49) .. controls (370.77,172.29) and (370.01,172.94) .. (369.08,172.94) .. controls (368.15,172.94) and (367.39,172.29) .. (367.39,171.49) -- cycle ;
%Shape: Ellipse [id:dp9654915371918213]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (370.18,177.12) .. controls (370.18,176.32) and (370.94,175.67) .. (371.87,175.67) .. controls (372.8,175.67) and (373.56,176.32) .. (373.56,177.12) .. controls (373.56,177.92) and (372.8,178.57) .. (371.87,178.57) .. controls (370.94,178.57) and (370.18,177.92) .. (370.18,177.12) -- cycle ;
%Shape: Ellipse [id:dp15713183567776468]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (373.53,183.35) .. controls (373.53,182.55) and (374.29,181.9) .. (375.22,181.9) .. controls (376.15,181.9) and (376.91,182.55) .. (376.91,183.35) .. controls (376.91,184.15) and (376.15,184.8) .. (375.22,184.8) .. controls (374.29,184.8) and (373.53,184.15) .. (373.53,183.35) -- cycle ;
%Shape: Ellipse [id:dp9239425899364824]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (377.52,190.32) .. controls (377.52,189.52) and (378.28,188.87) .. (379.21,188.87) .. controls (380.14,188.87) and (380.9,189.52) .. (380.9,190.32) .. controls (380.9,191.12) and (380.14,191.77) .. (379.21,191.77) .. controls (378.28,191.77) and (377.52,191.12) .. (377.52,190.32) -- cycle ;
%Shape: Ellipse [id:dp9110579702631929]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (381.35,196.88) .. controls (381.35,196.08) and (382.1,195.43) .. (383.04,195.43) .. controls (383.97,195.43) and (384.72,196.08) .. (384.72,196.88) .. controls (384.72,197.68) and (383.97,198.33) .. (383.04,198.33) .. controls (382.1,198.33) and (381.35,197.68) .. (381.35,196.88) -- cycle ;
%Shape: Ellipse [id:dp24310198218055246]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (385.67,204.67) .. controls (385.67,203.87) and (386.43,203.22) .. (387.36,203.22) .. controls (388.29,203.22) and (389.05,203.87) .. (389.05,204.67) .. controls (389.05,205.47) and (388.29,206.12) .. (387.36,206.12) .. controls (386.43,206.12) and (385.67,205.47) .. (385.67,204.67) -- cycle ;
%Shape: Ellipse [id:dp7388999956457485]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (381.21,171.49) .. controls (381.21,170.69) and (381.96,170.04) .. (382.9,170.04) .. controls (383.83,170.04) and (384.58,170.69) .. (384.58,171.49) .. controls (384.58,172.29) and (383.83,172.94) .. (382.9,172.94) .. controls (381.96,172.94) and (381.21,172.29) .. (381.21,171.49) -- cycle ;
%Shape: Ellipse [id:dp5640310139705571]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (384,177) .. controls (384,176.2) and (384.75,175.55) .. (385.69,175.55) .. controls (386.62,175.55) and (387.38,176.2) .. (387.38,177) .. controls (387.38,177.8) and (386.62,178.45) .. (385.69,178.45) .. controls (384.75,178.45) and (384,177.8) .. (384,177) -- cycle ;
%Shape: Ellipse [id:dp2695616969380312]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (387.63,183.23) .. controls (387.63,182.43) and (388.38,181.78) .. (389.32,181.78) .. controls (390.25,181.78) and (391,182.43) .. (391,183.23) .. controls (391,184.03) and (390.25,184.68) .. (389.32,184.68) .. controls (388.38,184.68) and (387.63,184.03) .. (387.63,183.23) -- cycle ;
%Shape: Ellipse [id:dp4126882239946579]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (391.49,190.24) .. controls (391.49,189.44) and (392.25,188.79) .. (393.18,188.79) .. controls (394.11,188.79) and (394.87,189.44) .. (394.87,190.24) .. controls (394.87,191.04) and (394.11,191.69) .. (393.18,191.69) .. controls (392.25,191.69) and (391.49,191.04) .. (391.49,190.24) -- cycle ;
%Shape: Ellipse [id:dp7640055339758918]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (395.16,197) .. controls (395.16,196.2) and (395.92,195.55) .. (396.85,195.55) .. controls (397.78,195.55) and (398.54,196.2) .. (398.54,197) .. controls (398.54,197.8) and (397.78,198.45) .. (396.85,198.45) .. controls (395.92,198.45) and (395.16,197.8) .. (395.16,197) -- cycle ;
%Shape: Ellipse [id:dp8721356938843838]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (399.35,204.79) .. controls (399.35,203.99) and (400.11,203.34) .. (401.04,203.34) .. controls (401.97,203.34) and (402.73,203.99) .. (402.73,204.79) .. controls (402.73,205.59) and (401.97,206.24) .. (401.04,206.24) .. controls (400.11,206.24) and (399.35,205.59) .. (399.35,204.79) -- cycle ;
%Straight Lines [id:da7979534893976028]
\draw [color={rgb, 255:red, 127; green, 117; blue, 117 }  ,draw opacity=1 ] [dash pattern={on 4.5pt off 4.5pt}]  (287.87,73.06) -- (287.87,183.38) ;
%Straight Lines [id:da03397341825975597]
\draw    (176.12,166.91) -- (148.03,213.62) ;
%Straight Lines [id:da855504113906919]
\draw    (148.03,213.62) -- (151.25,213.56) ;
%Straight Lines [id:da7263472575025891]
\draw    (176.12,166.91) -- (179.34,166.85) ;
%Straight Lines [id:da3871270427067002]
\draw    (420.92,213.12) -- (424.14,213.05) ;
%Straight Lines [id:da5505655212811509]
\draw    (393.8,167) -- (397.02,166.94) ;
%Straight Lines [id:da3454569240651897]
\draw    (397.02,166.94) -- (424.14,213.05) ;
%Straight Lines [id:da9496307475947867]
\draw [color={rgb, 255:red, 162; green, 157; blue, 157 }  ,draw opacity=1 ]   (159.76,153.9) -- (179.23,153.7) ;
\draw [shift={(182.23,153.66)}, rotate = 539.39] [fill={rgb, 255:red, 162; green, 157; blue, 157 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da7018774365527356]
\draw [color={rgb, 255:red, 162; green, 157; blue, 157 }  ,draw opacity=1 ]   (159.76,153.9) -- (150.65,164.89) ;
\draw [shift={(148.74,167.2)}, rotate = 309.65999999999997] [fill={rgb, 255:red, 162; green, 157; blue, 157 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Shape: Wave [id:dp8230250028099939]
\draw  [color={rgb, 255:red, 229; green, 21; blue, 21 }  ,draw opacity=1 ][line width=1.5]  (287.6,178.6) .. controls (302.2,174.36) and (316.1,170.3) .. (316.1,165.6) .. controls (316.1,160.9) and (302.2,156.84) .. (287.6,152.6) .. controls (273,148.36) and (259.1,144.3) .. (259.1,139.6) .. controls (259.1,134.9) and (273,130.84) .. (287.6,126.6) .. controls (302.2,122.36) and (316.1,118.3) .. (316.1,113.6) .. controls (316.1,108.9) and (302.2,104.84) .. (287.6,100.6) .. controls (273,96.36) and (259.1,92.3) .. (259.1,87.6) .. controls (259.1,83.15) and (271.51,79.29) .. (285.19,75.3) ;
%Straight Lines [id:da5686545133121326]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (287.8,80.6) -- (275.2,80.6) ;
\draw [shift={(272.2,80.6)}, rotate = 360] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da782823803042028]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (288.2,94.2) -- (275.2,94.53) ;
\draw [shift={(272.2,94.6)}, rotate = 358.57] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da871137874278286]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (288.6,87.2) -- (262.4,87.74) ;
\draw [shift={(259.4,87.8)}, rotate = 358.82] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da5553069086130265]
\draw [color={rgb, 255:red, 162; green, 157; blue, 157 }  ,draw opacity=1 ]   (159.76,153.9) -- (159.8,134) ;
\draw [shift={(159.8,131)}, rotate = 450.09] [fill={rgb, 255:red, 162; green, 157; blue, 157 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da6699653458590744]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (288.2,120.6) -- (299.6,120.6) ;
\draw [shift={(302.6,120.6)}, rotate = 180] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da2953745491779931]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (288.6,106.6) -- (301.6,106.6) ;
\draw [shift={(304.6,106.6)}, rotate = 180] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da7676872185404988]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (287.8,114) -- (312,113.47) ;
\draw [shift={(315,113.4)}, rotate = 538.74] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da6047601791922992]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (287.8,132.2) -- (275.2,132.2) ;
\draw [shift={(272.2,132.2)}, rotate = 360] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da08573309540969332]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (288.2,145.8) -- (275.2,146.13) ;
\draw [shift={(272.2,146.2)}, rotate = 358.57] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da0011618805281727873]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (288.6,138.8) -- (262.4,139.34) ;
\draw [shift={(259.4,139.4)}, rotate = 358.82] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da8026810749325723]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (287,172.6) -- (298.4,172.6) ;
\draw [shift={(301.4,172.6)}, rotate = 180] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da18748197647149523]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (287.4,158.6) -- (300.4,158.6) ;
\draw [shift={(303.4,158.6)}, rotate = 180] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da8245101205852268]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (286.6,166) -- (310.8,165.47) ;
\draw [shift={(313.8,165.4)}, rotate = 538.74] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da7887384290153587]
\draw [color={rgb, 255:red, 20; green, 88; blue, 145 }  ,draw opacity=1 ][line width=1.5]    (229.92,180.08) -- (229.78,134.23) ;
\draw [shift={(229.77,130.23)}, rotate = 449.82] [fill={rgb, 255:red, 20; green, 88; blue, 145 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (9.29,-4.46) -- (0,0) -- (9.29,4.46) -- cycle    ;
%Straight Lines [id:da8482623274180794]
\draw [color={rgb, 255:red, 20; green, 88; blue, 145 }  ,draw opacity=1 ][line width=1.5]    (250.23,180.08) -- (250.23,134.23) ;
\draw [shift={(250.23,130.23)}, rotate = 450] [fill={rgb, 255:red, 20; green, 88; blue, 145 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (9.29,-4.46) -- (0,0) -- (9.29,4.46) -- cycle    ;
%Straight Lines [id:da28732418115198444]
\draw [color={rgb, 255:red, 20; green, 88; blue, 145 }  ,draw opacity=1 ][line width=1.5]    (270.38,180.08) -- (270.24,134.38) ;
\draw [shift={(270.23,130.38)}, rotate = 449.82] [fill={rgb, 255:red, 20; green, 88; blue, 145 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (9.29,-4.46) -- (0,0) -- (9.29,4.46) -- cycle    ;
%Straight Lines [id:da23686823952523528]
\draw [color={rgb, 255:red, 20; green, 88; blue, 145 }  ,draw opacity=1 ][line width=1.5]    (290.38,179.92) -- (290.24,134.23) ;
\draw [shift={(290.23,130.23)}, rotate = 449.82] [fill={rgb, 255:red, 20; green, 88; blue, 145 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (9.29,-4.46) -- (0,0) -- (9.29,4.46) -- cycle    ;
%Straight Lines [id:da3497985465503395]
\draw [color={rgb, 255:red, 20; green, 88; blue, 145 }  ,draw opacity=1 ][line width=1.5]    (310.23,180.08) -- (310.09,134.38) ;
\draw [shift={(310.08,130.38)}, rotate = 449.82] [fill={rgb, 255:red, 20; green, 88; blue, 145 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (9.29,-4.46) -- (0,0) -- (9.29,4.46) -- cycle    ;
%Straight Lines [id:da4475685784611998]
\draw [color={rgb, 255:red, 20; green, 88; blue, 145 }  ,draw opacity=1 ][line width=1.5]    (330.08,180.23) -- (329.94,134.54) ;
\draw [shift={(329.92,130.54)}, rotate = 449.82] [fill={rgb, 255:red, 20; green, 88; blue, 145 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (9.29,-4.46) -- (0,0) -- (9.29,4.46) -- cycle    ;

% Text Node
\draw (134.77,155.14) node [anchor=north west][inner sep=0.75pt]  [font=\normalsize]  {$\mathbf{x}$};
% Text Node
\draw (183.93,136.77) node [anchor=north west][inner sep=0.75pt]  [font=\normalsize]  {$\mathbf{y}$};
% Text Node
\draw (145.06,72.2) node [anchor=north west][inner sep=0.75pt]  [font=\normalsize] [align=left] {\begin{minipage}[lt]{74.95912pt}\setlength\topsep{0pt}
\begin{center}
\textcolor[rgb]{0.82,0.01,0.11}{{\fontfamily{helvet}\selectfont Dressing Field's}}\\\textcolor[rgb]{0.82,0.01,0.11}{{\fontfamily{helvet}\selectfont Elctric Field}}
\end{center}

\end{minipage}};
% Text Node
\draw (145.93,116.37) node [anchor=north west][inner sep=0.75pt]  [font=\normalsize]  {$\mathbf{z}$};
% Text Node
\draw (347.46,140.94) node [anchor=north west][inner sep=0.75pt]  [font=\normalsize] [align=left] {\begin{minipage}[lt]{69.05196000000001pt}\setlength\topsep{0pt}
\begin{center}
\textcolor[rgb]{0.08,0.35,0.57}{{\fontfamily{helvet}\selectfont Magnetic Field}}
\end{center}

\end{minipage}};

\end{tikzpicture}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%COPY HERE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}