### 2021-04-09 theory/M2DEG.tex
Add scattering calculations

\documentclass[a4paper]{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=3cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{url}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{physics}
\usepackage{amssymb}
\usepackage{mathtools}

\numberwithin{equation}{subsection}
\newcommand{\mb}[1]{\mathbf{#1}}
\title{Magnetic propeties of a two dimentional electron gas strongly coupled to lights}
\author{K.Dini, O.V. Kibis and I.A. Shelykh}
\numberwithin{equation}{section}

\begin{document}

\maketitle

% \input{sec_01.tex}
% \newpage
\input{sec_02.tex}

\end{document}