### 2021-04-23 theory/figures/fig2.tex
Add figure 2 scattering

\documentclass{standalone}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tikz,pgf} %and any other packages or tikzlibraries your picture needs
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{physics}
\usepackage{amssymb}
\usepackage{mathtools}
\usetikzlibrary{shapes,arrows,patterns}

\begin{document}

% Pattern Info

\tikzset{
pattern size/.store in=\mcSize,
pattern size = 5pt,
pattern thickness/.store in=\mcThickness,
pattern thickness = 0.3pt,
pattern radius/.store in=\mcRadius,
pattern radius = 1pt}
\makeatletter
\pgfutil@ifundefined{pgf@pattern@name@_ahw2g2prc}{
\pgfdeclarepatternformonly[\mcThickness,\mcSize]{_ahw2g2prc}
{\pgfqpoint{0pt}{-\mcThickness}}
{\pgfpoint{\mcSize}{\mcSize}}
{\pgfpoint{\mcSize}{\mcSize}}
{
\pgfsetcolor{\tikz@pattern@color}
\pgfsetlinewidth{\mcThickness}
\pgfpathmoveto{\pgfqpoint{0pt}{\mcSize}}
\pgfpathlineto{\pgfpoint{\mcSize+\mcThickness}{-\mcThickness}}
\pgfusepath{stroke}
}}
\makeatother

% Pattern Info

\tikzset{
pattern size/.store in=\mcSize,
pattern size = 5pt,
pattern thickness/.store in=\mcThickness,
pattern thickness = 0.3pt,
pattern radius/.store in=\mcRadius,
pattern radius = 1pt}
\makeatletter
\pgfutil@ifundefined{pgf@pattern@name@_hzem6n8bo}{
\pgfdeclarepatternformonly[\mcThickness,\mcSize]{_hzem6n8bo}
{\pgfqpoint{0pt}{0pt}}
{\pgfpoint{\mcSize+\mcThickness}{\mcSize+\mcThickness}}
{\pgfpoint{\mcSize}{\mcSize}}
{
\pgfsetcolor{\tikz@pattern@color}
\pgfsetlinewidth{\mcThickness}
\pgfpathmoveto{\pgfqpoint{0pt}{0pt}}
\pgfpathlineto{\pgfpoint{\mcSize+\mcThickness}{\mcSize+\mcThickness}}
\pgfusepath{stroke}
}}
\makeatother

% Pattern Info

\tikzset{
pattern size/.store in=\mcSize,
pattern size = 5pt,
pattern thickness/.store in=\mcThickness,
pattern thickness = 0.3pt,
pattern radius/.store in=\mcRadius,
pattern radius = 1pt}
\makeatletter
\pgfutil@ifundefined{pgf@pattern@name@_bmptxkju1}{
\pgfdeclarepatternformonly[\mcThickness,\mcSize]{_bmptxkju1}
{\pgfqpoint{0pt}{0pt}}
{\pgfpoint{\mcSize+\mcThickness}{\mcSize+\mcThickness}}
{\pgfpoint{\mcSize}{\mcSize}}
{
\pgfsetcolor{\tikz@pattern@color}
\pgfsetlinewidth{\mcThickness}
\pgfpathmoveto{\pgfqpoint{0pt}{0pt}}
\pgfpathlineto{\pgfpoint{\mcSize+\mcThickness}{\mcSize+\mcThickness}}
\pgfusepath{stroke}
}}
\makeatother
\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300

%Shape: Rectangle [id:dp11071761337500274]
\draw  [color={rgb, 255:red, 255; green, 255; blue, 255 }  ,draw opacity=1 ][pattern=_ahw2g2prc,pattern size=5.550000000000001pt,pattern thickness=0.75pt,pattern radius=0pt, pattern color={rgb, 255:red, 198; green, 198; blue, 198}] (190.29,116.71) -- (409.57,116.71) -- (409.57,143.29) -- (190.29,143.29) -- cycle ;
%Straight Lines [id:da38313797800163174]
\draw [line width=1.5]    (300.14,196.43) -- (299.87,74.43) ;
\draw [shift={(299.86,70.43)}, rotate = 449.87] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (11.61,-5.58) -- (0,0) -- (11.61,5.58) -- cycle    ;
%Straight Lines [id:da2885312957054813]
\draw [line width=1.5]    (190.14,190.14) -- (410.43,190.43) ;
%Straight Lines [id:da014828325076878812]
\draw    (189.86,160.14) -- (409.57,159.86) ;
%Straight Lines [id:da030770486462505353]
\draw [color={rgb, 255:red, 74; green, 74; blue, 74 }  ,draw opacity=1 ]   (189.86,130.14) -- (409.86,130.14) ;
%Straight Lines [id:da1258041236737828]
\draw    (189.86,100.14) -- (409.86,100.14) ;
%Straight Lines [id:da549610540001811]
\draw [line width=0.75]  [dash pattern={on 4.5pt off 4.5pt}]  (190.29,115) -- (409.86,115.57) ;
%Straight Lines [id:da8170920135053421]
\draw  [dash pattern={on 4.5pt off 4.5pt}]  (190,144.71) -- (409.57,145.29) ;
%Straight Lines [id:da09440943310141292]
\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ]   (190.29,123.14) -- (410.14,123.57) ;
%Shape: Circle [id:dp2109394726683369]
\draw  [pattern=_hzem6n8bo,pattern size=1.9500000000000002pt,pattern thickness=0.75pt,pattern radius=0pt, pattern color={rgb, 255:red, 155; green, 155; blue, 155}] (326.29,130.21) .. controls (326.29,128.12) and (327.98,126.43) .. (330.07,126.43) .. controls (332.16,126.43) and (333.86,128.12) .. (333.86,130.21) .. controls (333.86,132.31) and (332.16,134) .. (330.07,134) .. controls (327.98,134) and (326.29,132.31) .. (326.29,130.21) -- cycle ;
%Shape: Circle [id:dp6488061807957903]
\draw  [pattern=_bmptxkju1,pattern size=1.9500000000000002pt,pattern thickness=0.75pt,pattern radius=0pt, pattern color={rgb, 255:red, 155; green, 155; blue, 155}] (356.29,123.36) .. controls (356.29,121.27) and (357.98,119.57) .. (360.07,119.57) .. controls (362.16,119.57) and (363.86,121.27) .. (363.86,123.36) .. controls (363.86,125.45) and (362.16,127.14) .. (360.07,127.14) .. controls (357.98,127.14) and (356.29,125.45) .. (356.29,123.36) -- cycle ;
%Curve Lines [id:da24190138502732017]
\draw [color={rgb, 255:red, 15; green, 77; blue, 150 }  ,draw opacity=1 ]   (329.86,125) .. controls (332.29,109.61) and (348.75,101.87) .. (356.82,115.78) ;
\draw [shift={(358.14,118.43)}, rotate = 246.91] [fill={rgb, 255:red, 15; green, 77; blue, 150 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (5.36,-2.57) -- (0,0) -- (5.36,2.57) -- (3.56,0) -- cycle    ;
%Straight Lines [id:da18098817640260245]
\draw    (330.14,185) -- (330.29,195.86) ;
%Straight Lines [id:da950065161038371]
\draw    (359.92,185.11) -- (360.06,195.97) ;
%Straight Lines [id:da6657355377016125]
\draw    (340.14,185) -- (340.29,195.86) ;
%Straight Lines [id:da588941077354121]
\draw    (350.14,185.57) -- (350.29,196.43) ;
%Straight Lines [id:da9026150195477192]
\draw    (370.14,184.56) -- (370.29,195.41) ;
%Straight Lines [id:da7248973636669813]
\draw    (399.92,184.67) -- (400.06,195.52) ;
%Straight Lines [id:da05943771455735014]
\draw    (380.14,184.56) -- (380.29,195.41) ;
%Straight Lines [id:da7726684154046342]
\draw    (390.14,185.13) -- (390.29,195.98) ;
%Straight Lines [id:da9150601062632764]
\draw    (250.14,185.22) -- (250.29,196.08) ;
%Straight Lines [id:da8099410498362345]
\draw    (279.92,185.33) -- (280.06,196.19) ;
%Straight Lines [id:da6236544524998151]
\draw    (260.14,185.22) -- (260.29,196.08) ;
%Straight Lines [id:da5921146606185723]
\draw    (270.14,185.79) -- (270.29,196.65) ;
%Straight Lines [id:da5991996011400991]
\draw    (290.14,184.78) -- (290.29,195.63) ;
%Straight Lines [id:da6482455334673505]
\draw    (319.92,184.89) -- (320.06,195.75) ;
%Straight Lines [id:da871163669545836]
\draw    (310.14,185.35) -- (310.29,196.21) ;
%Straight Lines [id:da6483925428035442]
\draw    (199.48,184.67) -- (199.62,195.52) ;
%Straight Lines [id:da685643393265353]
\draw    (209.7,184.11) -- (209.84,194.97) ;
%Straight Lines [id:da7838129922931183]
\draw    (239.48,184.22) -- (239.62,195.08) ;
%Straight Lines [id:da7719533009206734]
\draw    (219.7,184.11) -- (219.84,194.97) ;
%Straight Lines [id:da2831023061884146]
\draw    (229.7,184.68) -- (229.84,195.54) ;

% Text Node
\draw (299.71,51.54) node [anchor=north west][inner sep=0.75pt]    {$Energy$};
% Text Node
\draw (415.71,180.69) node [anchor=north west][inner sep=0.75pt]    {$k_{x}$};
% Text Node
\draw (411.86,125.97) node [anchor=north west][inner sep=0.75pt]  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,opacity=1 ]  {$\varepsilon_{\beta} =\varepsilon_{n_{\beta}}$};
% Text Node
\draw (176.29,117.11) node [anchor=north west][inner sep=0.75pt]  [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,opacity=1 ]  {$\varepsilon$};
% Text Node
\draw (308,132.105) node [anchor=north west][inner sep=0.75pt]  [font=\tiny]  {$\ket{\psi_{\beta}}$};
% Text Node
\draw (360.29,104.4) node [anchor=north west][inner sep=0.75pt]  [font=\tiny]  {$\ket{\Psi_{\alpha}}$};

\end{tikzpicture}

\end{document}