### 2021-05-18 theory/sec_08.tex
Add conductivity analysis

\section{Manipulate Conductivity in Quantum Hall System}

Now using the relations derived in Eq. \eqref{5.34} and Eq. \eqref{7.17} we can derive that
\begin{equation} \label{8.1}
  \Gamma_n = \qty(\frac{\hbar }{2\tau(\varepsilon_n)}) =
  \frac{1}{2}\Gamma^{00}_{N}(\tilde{I})
\end{equation}
and this will be
\begin{equation} \label{8.2}
  \Gamma_n(\tilde{I}) = 0.12\times
  \qty[
  \frac
  {\int_{-\infty}^{\infty} d {k}_1 \;
  J_0^2\qty(2.090\sqrt{\tilde{I}}\times{k}_1)
  \qty|
  \int_{-\infty}^{\infty} d{k}_2 \;
  \tilde{\chi}_{N}\qty(2.342 \times k_2)
  \tilde{\chi}_{N}\qty(2.342 \times \qty[{k}_1 - {k}_2])|^2}
  {\int_{-\infty}^{\infty} d {k}_1 \;
  \qty|
  \int_{-\infty}^{\infty} d{k}_2 \;
  \tilde{\chi}_{0}\qty(2.342 \times k_2)
  \tilde{\chi}_{0}\qty(2.342 \times \qty[{k}_1 - {k}_2])|^2}
  ]^{1/2} \text{meV}
\end{equation}

\noindent
In addition we can calculate the cyclotron energy as
\begin{equation} \label{8.3}
  \hbar\omega_0 = 1.95663 \;\text{meV}
\end{equation}
and
\begin{equation} \label{8.4}
  \gamma_n = \frac{\Gamma_n}{\hbar \omega_0} = 0.06133 \times \Lambda_n(\tilde{I}) \approx 0.061 \Lambda_n(\tilde{I})
\end{equation}

\noindent
Now we can use this into our conductivity expression derived in Eq. \eqref{7.23} and present the normalized transverse conductivity as a function of fermi energy and intensity of the dressing field
\begin{equation} \label{8.5}
  \begin{aligned}
    \sigma^{xx}(X_F,\tilde{I}) & =
    \sum_{n}
    \frac{\qty(n+1)}{0.0037\Lambda_n \Lambda_{n+1}}
    \qty[
      \frac{1}
      {
        1 + \qty(\frac{X_F - n -1}{0.06\Lambda_n})^2
      }
    ]
    \qty[
      \frac{1}
      {
        1 + \qty(\frac{X_F - n}{0.06\Lambda_{n+1}})^2
      }
    ]
  \end{aligned}
\end{equation}

\hfill$\blacksquare$