### 2021-05-19 theory/sec_08.tex
Add calculated normalized values

\section{Manipulate Conductivity in Quantum Hall System}

Now using the relations derived in Eq. \eqref{5.34} and Eq. \eqref{7.17} we can derive that
\begin{equation} \label{8.1}
  \Gamma_n = \qty(\frac{\hbar }{2\tau(\varepsilon_n)}) =
  \frac{1}{2}\Gamma^{00}_{N}(\tilde{I})
\end{equation}
and this will be
\begin{equation} \label{8.2}
  \Gamma_n(\tilde{I}) = 0.12\times
  \qty[
  \frac
  {\int_{-\infty}^{\infty} d {k}_1 \;
  J_0^2\qty(2.090\sqrt{\tilde{I}}\times{k}_1)
  \qty|
  \int_{-\infty}^{\infty} d{k}_2 \;
  \tilde{\chi}_{N}\qty(2.342 \times k_2)
  \tilde{\chi}_{N}\qty(2.342 \times \qty[{k}_1 - {k}_2])|^2}
  {\int_{-\infty}^{\infty} d {k}_1 \;
  \qty|
  \int_{-\infty}^{\infty} d{k}_2 \;
  \tilde{\chi}_{0}\qty(2.342 \times k_2)
  \tilde{\chi}_{0}\qty(2.342 \times \qty[{k}_1 - {k}_2])|^2}
  ]^{1/2} \text{meV}
\end{equation}

\noindent
In addition we can calculate the cyclotron energy as
\begin{equation} \label{8.3}
  \hbar\omega_0 = 1.95663 \;\text{meV}
\end{equation}
and
\begin{equation} \label{8.4}
  \gamma_n = \frac{\Gamma_n}{\hbar \omega_0} = 0.06133 \times \Lambda_n(\tilde{I}) \approx 0.061 \Lambda_n(\tilde{I})
\end{equation}

\noindent
Now we can use this into our conductivity expression derived in Eq. \eqref{7.23} and present the normalized transverse conductivity as a function of fermi energy and intensity of the dressing field
\begin{equation} \label{8.5}
  \begin{aligned}
    \sigma^{xx}(X_F,\tilde{I}) & =
    \sum_{n}
    \frac{\qty(n+1)}{0.0037\Lambda_n \Lambda_{n+1}}
    \qty[
      \frac{1}
      {
        1 + \qty(\frac{X_F - n -1}{0.06\Lambda_n})^2
      }
    ]
    \qty[
      \frac{1}
      {
        1 + \qty(\frac{X_F - n}{0.06\Lambda_{n+1}})^2
      }
    ]
  \end{aligned}
\end{equation}
where
\begin{equation} \label{8.6}
    \Lambda_n  =
    \qty[
    \frac
    {\int_{-\infty}^{\infty} d {k}_1 \;
    J_0^2\qty(2.090\sqrt{\tilde{I}}\times{k}_1)
    \qty|
    \int_{-\infty}^{\infty} d{k}_2 \;
    \tilde{\chi}_{n}\qty(2.342 \times k_2)
    \tilde{\chi}_{n}\qty(2.342 \times \qty[{k}_1 - {k}_2])|^2}
    {\int_{-\infty}^{\infty} d {k}_1 \;
    \qty|
    \int_{-\infty}^{\infty} d{k}_2 \;
    \tilde{\chi}_{0}\qty(2.342 \times k_2)
    \tilde{\chi}_{0}\qty(2.342 \times \qty[{k}_1 - {k}_2])|^2}
    ]^{1/2}.
\end{equation}
\hfill$\blacksquare$

\noindent
Now using Eq. \eqref{8.6} we can calculate the normalized broadening of each Landau levels as given in Table \ref{tab:8.1}.

\begin{table}[ht!]
\begin{center}
\begin{tabular}{ |c|c|c|c|c|c|c|c|c|c|c|c| }
 \hline
 ${\tilde{I}}$
 & $n=0$ & $n=1$ & $n=2$ & $n=3$ & $n=4$ & $n=5$ & $n=6$ & $n=7$ & $n=8$ & $n=9$
 & $n=10$
 \\ [0.5ex] \hline\hline
 0  & $1.0000$ & $0.8660$ & $0.8004$ & $0.7578$ & $0.7266$
 & $0.7021$ & $0.6821$ & $0.6653$ & $0.6507$ & $0.6380$& $0.6267$ \\ \hline
 1  & $0.8546$ & $0.7037$ & $0.6502$ & $0.6114$ & $0.5810$
 & $0.5584$ & $0.5416$ & $0.5280$ & $0.5160$ & $0.5050$ & $0.4948$ \\ \hline
 4  & $0.6875$ & $0.6345$ & $0.5902$ & $0.5528$ & $0.5307$
 & $0.5118$ & $0.4958$ & $0.4840$ & $0.4730$ & $0.4629$  & $0.4547$ \\ \hline
 9  & $0.5936$ & $0.5539$ & $0.5333$ & $0.5180$ & $0.5006$
 & $0.4812$ & $0.4685$ & $0.4564$ & $0.4469$ & $0.4377$  & $0.4305$ \\ \hline
\end{tabular}
\caption {\label{tab:8.1} $\Lambda_n$ values for each Landau level by changing  applied dressing field intensity ($\tilde{I}$). }
\end{center}
\end{table}

\hfill$\blacksquare$