### 2021-05-21 theory/M2DEG.tex

\documentclass[a4paper]{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=3cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{graphicx}
\usepackage{url}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{physics}
\usepackage{amssymb}
\usepackage{mathtools}

\numberwithin{equation}{subsection}
\newcommand{\mb}[1]{\mathbf{#1}}
\title{Floquet-Drude Conductivity in Dressed Quantum Hall Systems}
\author{Kosala Sananthana Herath}
\numberwithin{equation}{section}

\begin{document}

\maketitle

\input{sec_01.tex}
\newpage
% \input{sec_02.tex}
% \newpage
\input{sec_03.tex}
\newpage
\input{sec_04.tex}
\newpage
\input{sec_05.tex}
\newpage
\input{sec_06.tex}
\newpage
\input{sec_07.tex}
\newpage
\input{sec_08.tex}

\end{document}