### 2021-06-28 theory/figures/fig1.tex
Add section 1 of paper

\documentclass{standalone}

\usepackage{tikz,pgf} %and any other packages or tikzlibraries your picture needs

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%COPY HERE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300

%Straight Lines [id:da5726274877822322]
\draw [color={rgb, 255:red, 127; green, 117; blue, 117 }  ,draw opacity=1 ] [dash pattern={on 4.5pt off 4.5pt}]  (309.51,73.75) -- (309.87,203.38) ;
\draw [shift={(309.5,70.75)}, rotate = 89.84] [fill={rgb, 255:red, 127; green, 117; blue, 117 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (5.36,-2.57) -- (0,0) -- (5.36,2.57) -- cycle    ;
%Straight Lines [id:da33511458189374754]
\draw    (171.25,233.56) -- (440.92,233.12) ;
%Straight Lines [id:da3398276064046053]
\draw    (198.99,187.4) -- (413.4,187.4) ;
%Straight Lines [id:da7839634174109911]
\draw    (199.39,187) -- (171.25,233.56) ;
%Straight Lines [id:da05894655796955206]
\draw    (413.8,187) -- (440.92,233.12) ;
%Straight Lines [id:da13536583120363344]
\draw    (306.6,187) -- (306.08,233.34) ;
%Straight Lines [id:da4997587586993617]
\draw    (293.92,186.96) -- (290.68,233.56) ;
%Straight Lines [id:da9722317036710604]
\draw    (279.41,186.96) -- (273.18,233.3) ;
%Straight Lines [id:da026737631204194257]
\draw    (265.3,186.91) -- (257.52,233.3) ;
%Straight Lines [id:da1602511216081195]
\draw    (252.1,187.18) -- (240.33,233.82) ;
%Straight Lines [id:da1955990930085778]
\draw    (237.96,187.2) -- (221.3,233.56) ;
%Straight Lines [id:da21157278737030283]
\draw    (225.7,187.18) -- (205.02,233.82) ;
%Straight Lines [id:da7497787157955642]
\draw    (212.98,187.2) -- (189.98,233.82) ;
%Straight Lines [id:da8978850152403992]
\draw    (318.62,186.91) -- (322.61,233.56) ;
%Straight Lines [id:da8571932918195948]
\draw    (333.18,187.08) -- (341.21,233.92) ;
%Straight Lines [id:da28540902944632984]
\draw    (347.48,186.91) -- (359.76,233.3) ;
%Straight Lines [id:da9908810182455174]
\draw    (360.68,187.18) -- (377.57,233.82) ;
%Straight Lines [id:da9097146225861146]
\draw    (374.8,187.18) -- (394.76,233.56) ;
%Straight Lines [id:da7912008612279866]
\draw    (386.16,187.08) -- (412.26,233.56) ;
%Straight Lines [id:da5553223696891176]
\draw    (400.59,187.18) -- (425.77,233.3) ;
%Straight Lines [id:da4854495790277076]
\draw    (175.55,225.39) -- (436.51,224.6) ;
%Straight Lines [id:da8466426333509127]
\draw    (180.77,217.48) -- (431.6,216.96) ;
%Straight Lines [id:da11285301674845338]
\draw    (185.32,210.28) -- (427.36,210.06) ;
%Straight Lines [id:da6415671559000329]
\draw    (189.06,203.52) -- (423.31,203.25) ;
%Straight Lines [id:da09765847114278392]
\draw    (192.65,197.85) -- (419.54,197.33) ;
%Straight Lines [id:da9654913395744491]
\draw    (195.41,192.58) -- (416.16,191.79) ;
%Shape: Ellipse [id:dp5486828293018222]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (192.44,225.39) .. controls (192.44,224.59) and (193.19,223.94) .. (194.12,223.94) .. controls (195.06,223.94) and (195.81,224.59) .. (195.81,225.39) .. controls (195.81,226.19) and (195.06,226.84) .. (194.12,226.84) .. controls (193.19,226.84) and (192.44,226.19) .. (192.44,225.39) -- cycle ;
%Shape: Ellipse [id:dp36608685875382685]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (196.2,217.36) .. controls (196.2,216.56) and (196.96,215.91) .. (197.89,215.91) .. controls (198.82,215.91) and (199.58,216.56) .. (199.58,217.36) .. controls (199.58,218.16) and (198.82,218.81) .. (197.89,218.81) .. controls (196.96,218.81) and (196.2,218.16) .. (196.2,217.36) -- cycle ;
%Shape: Ellipse [id:dp5251439025998383]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (199.83,209.94) .. controls (199.83,209.14) and (200.59,208.49) .. (201.52,208.49) .. controls (202.45,208.49) and (203.21,209.14) .. (203.21,209.94) .. controls (203.21,210.74) and (202.45,211.39) .. (201.52,211.39) .. controls (200.59,211.39) and (199.83,210.74) .. (199.83,209.94) -- cycle ;
%Shape: Ellipse [id:dp4067558888223195]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (202.9,203.59) .. controls (202.9,202.79) and (203.66,202.14) .. (204.59,202.14) .. controls (205.52,202.14) and (206.28,202.79) .. (206.28,203.59) .. controls (206.28,204.39) and (205.52,205.04) .. (204.59,205.04) .. controls (203.66,205.04) and (202.9,204.39) .. (202.9,203.59) -- cycle ;
%Shape: Ellipse [id:dp3570254692613275]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (206.11,197.24) .. controls (206.11,196.44) and (206.87,195.79) .. (207.8,195.79) .. controls (208.73,195.79) and (209.49,196.44) .. (209.49,197.24) .. controls (209.49,198.04) and (208.73,198.69) .. (207.8,198.69) .. controls (206.87,198.69) and (206.11,198.04) .. (206.11,197.24) -- cycle ;
%Shape: Ellipse [id:dp30798846372533184]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (209.04,191.85) .. controls (209.04,191.05) and (209.8,190.4) .. (210.73,190.4) .. controls (211.66,190.4) and (212.42,191.05) .. (212.42,191.85) .. controls (212.42,192.65) and (211.66,193.3) .. (210.73,193.3) .. controls (209.8,193.3) and (209.04,192.65) .. (209.04,191.85) -- cycle ;
%Shape: Ellipse [id:dp1418023342038397]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (206.95,225.15) .. controls (206.95,224.35) and (207.71,223.7) .. (208.64,223.7) .. controls (209.57,223.7) and (210.33,224.35) .. (210.33,225.15) .. controls (210.33,225.95) and (209.57,226.6) .. (208.64,226.6) .. controls (207.71,226.6) and (206.95,225.95) .. (206.95,225.15) -- cycle ;
%Shape: Ellipse [id:dp32047691597756267]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (210.44,217.48) .. controls (210.44,216.68) and (211.19,216.03) .. (212.13,216.03) .. controls (213.06,216.03) and (213.82,216.68) .. (213.82,217.48) .. controls (213.82,218.28) and (213.06,218.93) .. (212.13,218.93) .. controls (211.19,218.93) and (210.44,218.28) .. (210.44,217.48) -- cycle ;
%Shape: Ellipse [id:dp49839354018590987]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (213.67,210.5) .. controls (213.67,209.7) and (214.43,209.05) .. (215.36,209.05) .. controls (216.29,209.05) and (217.05,209.7) .. (217.05,210.5) .. controls (217.05,211.3) and (216.29,211.95) .. (215.36,211.95) .. controls (214.43,211.95) and (213.67,211.3) .. (213.67,210.5) -- cycle ;
%Shape: Ellipse [id:dp8228751417402214]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (216.58,203.47) .. controls (216.58,202.67) and (217.33,202.02) .. (218.27,202.02) .. controls (219.2,202.02) and (219.96,202.67) .. (219.96,203.47) .. controls (219.96,204.27) and (219.2,204.92) .. (218.27,204.92) .. controls (217.33,204.92) and (216.58,204.27) .. (216.58,203.47) -- cycle ;
%Shape: Ellipse [id:dp06953462817928524]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (219.51,197.48) .. controls (219.51,196.68) and (220.27,196.03) .. (221.2,196.03) .. controls (222.13,196.03) and (222.89,196.68) .. (222.89,197.48) .. controls (222.89,198.28) and (222.13,198.93) .. (221.2,198.93) .. controls (220.27,198.93) and (219.51,198.28) .. (219.51,197.48) -- cycle ;
%Shape: Ellipse [id:dp3903462442234402]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (221.6,191.85) .. controls (221.6,191.05) and (222.36,190.4) .. (223.29,190.4) .. controls (224.22,190.4) and (224.98,191.05) .. (224.98,191.85) .. controls (224.98,192.65) and (224.22,193.3) .. (223.29,193.3) .. controls (222.36,193.3) and (221.6,192.65) .. (221.6,191.85) -- cycle ;
%Shape: Ellipse [id:dp4379467974609894]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (222.72,225.27) .. controls (222.72,224.47) and (223.47,223.82) .. (224.41,223.82) .. controls (225.34,223.82) and (226.1,224.47) .. (226.1,225.27) .. controls (226.1,226.07) and (225.34,226.72) .. (224.41,226.72) .. controls (223.47,226.72) and (222.72,226.07) .. (222.72,225.27) -- cycle ;
%Shape: Ellipse [id:dp26193777518096795]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (225.51,217.36) .. controls (225.51,216.56) and (226.27,215.91) .. (227.2,215.91) .. controls (228.13,215.91) and (228.89,216.56) .. (228.89,217.36) .. controls (228.89,218.16) and (228.13,218.81) .. (227.2,218.81) .. controls (226.27,218.81) and (225.51,218.16) .. (225.51,217.36) -- cycle ;
%Shape: Ellipse [id:dp019861246471317262]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (227.95,210.5) .. controls (227.95,209.7) and (228.7,209.05) .. (229.64,209.05) .. controls (230.57,209.05) and (231.32,209.7) .. (231.32,210.5) .. controls (231.32,211.3) and (230.57,211.95) .. (229.64,211.95) .. controls (228.7,211.95) and (227.95,211.3) .. (227.95,210.5) -- cycle ;
%Shape: Ellipse [id:dp1381162876501707]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (230.53,203.35) .. controls (230.53,202.55) and (231.29,201.9) .. (232.22,201.9) .. controls (233.15,201.9) and (233.91,202.55) .. (233.91,203.35) .. controls (233.91,204.15) and (233.15,204.8) .. (232.22,204.8) .. controls (231.29,204.8) and (230.53,204.15) .. (230.53,203.35) -- cycle ;
%Shape: Ellipse [id:dp5428095668243373]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (232.63,197.12) .. controls (232.63,196.32) and (233.38,195.67) .. (234.32,195.67) .. controls (235.25,195.67) and (236,196.32) .. (236,197.12) .. controls (236,197.92) and (235.25,198.57) .. (234.32,198.57) .. controls (233.38,198.57) and (232.63,197.92) .. (232.63,197.12) -- cycle ;
%Shape: Ellipse [id:dp18211764432101352]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (234.58,191.73) .. controls (234.58,190.93) and (235.34,190.28) .. (236.27,190.28) .. controls (237.2,190.28) and (237.96,190.93) .. (237.96,191.73) .. controls (237.96,192.53) and (237.2,193.18) .. (236.27,193.18) .. controls (235.34,193.18) and (234.58,192.53) .. (234.58,191.73) -- cycle ;
%Shape: Ellipse [id:dp75878279205988]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (240.72,225.39) .. controls (240.72,224.59) and (241.48,223.94) .. (242.41,223.94) .. controls (243.34,223.94) and (244.1,224.59) .. (244.1,225.39) .. controls (244.1,226.19) and (243.34,226.84) .. (242.41,226.84) .. controls (241.48,226.84) and (240.72,226.19) .. (240.72,225.39) -- cycle ;
%Shape: Ellipse [id:dp4571990490264608]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (242.81,217.48) .. controls (242.81,216.68) and (243.57,216.03) .. (244.5,216.03) .. controls (245.44,216.03) and (246.19,216.68) .. (246.19,217.48) .. controls (246.19,218.28) and (245.44,218.93) .. (244.5,218.93) .. controls (243.57,218.93) and (242.81,218.28) .. (242.81,217.48) -- cycle ;
%Shape: Ellipse [id:dp4013248465635957]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (244.53,210.5) .. controls (244.53,209.7) and (245.28,209.05) .. (246.21,209.05) .. controls (247.15,209.05) and (247.9,209.7) .. (247.9,210.5) .. controls (247.9,211.3) and (247.15,211.95) .. (246.21,211.95) .. controls (245.28,211.95) and (244.53,211.3) .. (244.53,210.5) -- cycle ;
%Shape: Ellipse [id:dp7663905432089348]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (246.16,203.59) .. controls (246.16,202.79) and (246.92,202.14) .. (247.85,202.14) .. controls (248.78,202.14) and (249.54,202.79) .. (249.54,203.59) .. controls (249.54,204.39) and (248.78,205.04) .. (247.85,205.04) .. controls (246.92,205.04) and (246.16,204.39) .. (246.16,203.59) -- cycle ;
%Shape: Ellipse [id:dp37843272203409817]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (247.7,197.36) .. controls (247.7,196.56) and (248.45,195.91) .. (249.39,195.91) .. controls (250.32,195.91) and (251.08,196.56) .. (251.08,197.36) .. controls (251.08,198.16) and (250.32,198.81) .. (249.39,198.81) .. controls (248.45,198.81) and (247.7,198.16) .. (247.7,197.36) -- cycle ;
%Shape: Ellipse [id:dp6447080382369588]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (249.09,191.97) .. controls (249.09,191.17) and (249.85,190.52) .. (250.78,190.52) .. controls (251.72,190.52) and (252.47,191.17) .. (252.47,191.97) .. controls (252.47,192.77) and (251.72,193.42) .. (250.78,193.42) .. controls (249.85,193.42) and (249.09,192.77) .. (249.09,191.97) -- cycle ;
%Shape: Ellipse [id:dp1815955543825718]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (262.77,191.85) .. controls (262.77,191.05) and (263.53,190.4) .. (264.46,190.4) .. controls (265.39,190.4) and (266.15,191.05) .. (266.15,191.85) .. controls (266.15,192.65) and (265.39,193.3) .. (264.46,193.3) .. controls (263.53,193.3) and (262.77,192.65) .. (262.77,191.85) -- cycle ;
%Shape: Ellipse [id:dp8880224712841644]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (261.79,197.48) .. controls (261.79,196.68) and (262.55,196.03) .. (263.48,196.03) .. controls (264.41,196.03) and (265.17,196.68) .. (265.17,197.48) .. controls (265.17,198.28) and (264.41,198.93) .. (263.48,198.93) .. controls (262.55,198.93) and (261.79,198.28) .. (261.79,197.48) -- cycle ;
%Shape: Ellipse [id:dp4989200469883939]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (260.68,203.23) .. controls (260.68,202.43) and (261.43,201.78) .. (262.37,201.78) .. controls (263.3,201.78) and (264.05,202.43) .. (264.05,203.23) .. controls (264.05,204.03) and (263.3,204.68) .. (262.37,204.68) .. controls (261.43,204.68) and (260.68,204.03) .. (260.68,203.23) -- cycle ;
%Shape: Ellipse [id:dp28395245427438764]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (259.72,210.1) .. controls (259.72,209.3) and (260.48,208.65) .. (261.41,208.65) .. controls (262.34,208.65) and (263.1,209.3) .. (263.1,210.1) .. controls (263.1,210.9) and (262.34,211.55) .. (261.41,211.55) .. controls (260.48,211.55) and (259.72,210.9) .. (259.72,210.1) -- cycle ;
%Shape: Ellipse [id:dp03156593400590513]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (258.44,217.36) .. controls (258.44,216.56) and (259.2,215.91) .. (260.13,215.91) .. controls (261.07,215.91) and (261.82,216.56) .. (261.82,217.36) .. controls (261.82,218.16) and (261.07,218.81) .. (260.13,218.81) .. controls (259.2,218.81) and (258.44,218.16) .. (258.44,217.36) -- cycle ;
%Shape: Ellipse [id:dp6428056825741797]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (257.19,225.03) .. controls (257.19,224.23) and (257.94,223.58) .. (258.88,223.58) .. controls (259.81,223.58) and (260.57,224.23) .. (260.57,225.03) .. controls (260.57,225.83) and (259.81,226.48) .. (258.88,226.48) .. controls (257.94,226.48) and (257.19,225.83) .. (257.19,225.03) -- cycle ;
%Shape: Ellipse [id:dp1475873632921516]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (277.01,191.85) .. controls (277.01,191.05) and (277.76,190.4) .. (278.69,190.4) .. controls (279.63,190.4) and (280.38,191.05) .. (280.38,191.85) .. controls (280.38,192.65) and (279.63,193.3) .. (278.69,193.3) .. controls (277.76,193.3) and (277.01,192.65) .. (277.01,191.85) -- cycle ;
%Shape: Ellipse [id:dp4429598972841515]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (276.17,197.36) .. controls (276.17,196.56) and (276.92,195.91) .. (277.86,195.91) .. controls (278.79,195.91) and (279.54,196.56) .. (279.54,197.36) .. controls (279.54,198.16) and (278.79,198.81) .. (277.86,198.81) .. controls (276.92,198.81) and (276.17,198.16) .. (276.17,197.36) -- cycle ;
%Shape: Ellipse [id:dp6989263176245077]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (275.33,203.47) .. controls (275.33,202.67) and (276.09,202.02) .. (277.02,202.02) .. controls (277.95,202.02) and (278.71,202.67) .. (278.71,203.47) .. controls (278.71,204.27) and (277.95,204.92) .. (277.02,204.92) .. controls (276.09,204.92) and (275.33,204.27) .. (275.33,203.47) -- cycle ;
%Shape: Ellipse [id:dp5884337821826009]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (274.56,210.37) .. controls (274.56,209.57) and (275.32,208.92) .. (276.25,208.92) .. controls (277.18,208.92) and (277.94,209.57) .. (277.94,210.37) .. controls (277.94,211.17) and (277.18,211.82) .. (276.25,211.82) .. controls (275.32,211.82) and (274.56,211.17) .. (274.56,210.37) -- cycle ;
%Shape: Ellipse [id:dp5456914529153225]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (273.52,217.24) .. controls (273.52,216.44) and (274.27,215.79) .. (275.2,215.79) .. controls (276.14,215.79) and (276.89,216.44) .. (276.89,217.24) .. controls (276.89,218.04) and (276.14,218.69) .. (275.2,218.69) .. controls (274.27,218.69) and (273.52,218.04) .. (273.52,217.24) -- cycle ;
%Shape: Ellipse [id:dp10200774997945894]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (272.4,225.15) .. controls (272.4,224.35) and (273.16,223.7) .. (274.09,223.7) .. controls (275.02,223.7) and (275.78,224.35) .. (275.78,225.15) .. controls (275.78,225.95) and (275.02,226.6) .. (274.09,226.6) .. controls (273.16,226.6) and (272.4,225.95) .. (272.4,225.15) -- cycle ;
%Shape: Ellipse [id:dp09546155197155715]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (291.94,191.85) .. controls (291.94,191.05) and (292.69,190.4) .. (293.63,190.4) .. controls (294.56,190.4) and (295.31,191.05) .. (295.31,191.85) .. controls (295.31,192.65) and (294.56,193.3) .. (293.63,193.3) .. controls (292.69,193.3) and (291.94,192.65) .. (291.94,191.85) -- cycle ;
%Shape: Ellipse [id:dp8888091900844011]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (291.52,197.24) .. controls (291.52,196.44) and (292.27,195.79) .. (293.21,195.79) .. controls (294.14,195.79) and (294.9,196.44) .. (294.9,197.24) .. controls (294.9,198.04) and (294.14,198.69) .. (293.21,198.69) .. controls (292.27,198.69) and (291.52,198.04) .. (291.52,197.24) -- cycle ;
%Shape: Ellipse [id:dp6682650227662896]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (291.24,203.23) .. controls (291.24,202.43) and (292,201.78) .. (292.93,201.78) .. controls (293.86,201.78) and (294.62,202.43) .. (294.62,203.23) .. controls (294.62,204.03) and (293.86,204.68) .. (292.93,204.68) .. controls (292,204.68) and (291.24,204.03) .. (291.24,203.23) -- cycle ;
%Shape: Ellipse [id:dp9294532115656315]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (290.68,210.5) .. controls (290.68,209.7) and (291.44,209.05) .. (292.37,209.05) .. controls (293.3,209.05) and (294.06,209.7) .. (294.06,210.5) .. controls (294.06,211.3) and (293.3,211.95) .. (292.37,211.95) .. controls (291.44,211.95) and (290.68,211.3) .. (290.68,210.5) -- cycle ;
%Shape: Ellipse [id:dp5568966606864136]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (290.12,217.12) .. controls (290.12,216.32) and (290.88,215.67) .. (291.81,215.67) .. controls (292.74,215.67) and (293.5,216.32) .. (293.5,217.12) .. controls (293.5,217.92) and (292.74,218.57) .. (291.81,218.57) .. controls (290.88,218.57) and (290.12,217.92) .. (290.12,217.12) -- cycle ;
%Shape: Ellipse [id:dp2121643224676708]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (289.7,225.03) .. controls (289.7,224.23) and (290.46,223.58) .. (291.39,223.58) .. controls (292.33,223.58) and (293.08,224.23) .. (293.08,225.03) .. controls (293.08,225.83) and (292.33,226.48) .. (291.39,226.48) .. controls (290.46,226.48) and (289.7,225.83) .. (289.7,225.03) -- cycle ;
%Shape: Ellipse [id:dp9886700240195399]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (304.5,191.79) .. controls (304.5,190.99) and (305.25,190.34) .. (306.19,190.34) .. controls (307.12,190.34) and (307.87,190.99) .. (307.87,191.79) .. controls (307.87,192.59) and (307.12,193.24) .. (306.19,193.24) .. controls (305.25,193.24) and (304.5,192.59) .. (304.5,191.79) -- cycle ;
%Shape: Ellipse [id:dp8920179824675918]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (304.8,197.19) .. controls (304.8,196.39) and (305.56,195.74) .. (306.49,195.74) .. controls (307.43,195.74) and (308.18,196.39) .. (308.18,197.19) .. controls (308.18,197.99) and (307.43,198.64) .. (306.49,198.64) .. controls (305.56,198.64) and (304.8,197.99) .. (304.8,197.19) -- cycle ;
%Shape: Ellipse [id:dp2107726543746229]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (304.5,203.38) .. controls (304.5,202.58) and (305.25,201.93) .. (306.19,201.93) .. controls (307.12,201.93) and (307.87,202.58) .. (307.87,203.38) .. controls (307.87,204.18) and (307.12,204.83) .. (306.19,204.83) .. controls (305.25,204.83) and (304.5,204.18) .. (304.5,203.38) -- cycle ;
%Shape: Ellipse [id:dp7703015900268393]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (304.65,210.17) .. controls (304.65,209.37) and (305.41,208.72) .. (306.34,208.72) .. controls (307.27,208.72) and (308.03,209.37) .. (308.03,210.17) .. controls (308.03,210.97) and (307.27,211.62) .. (306.34,211.62) .. controls (305.41,211.62) and (304.65,210.97) .. (304.65,210.17) -- cycle ;
%Shape: Ellipse [id:dp012791866206312008]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (304.5,217.22) .. controls (304.5,216.42) and (305.25,215.77) .. (306.19,215.77) .. controls (307.12,215.77) and (307.87,216.42) .. (307.87,217.22) .. controls (307.87,218.02) and (307.12,218.67) .. (306.19,218.67) .. controls (305.25,218.67) and (304.5,218.02) .. (304.5,217.22) -- cycle ;
%Shape: Ellipse [id:dp5121947850193356]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (304.34,224.99) .. controls (304.34,224.19) and (305.1,223.54) .. (306.03,223.54) .. controls (306.96,223.54) and (307.72,224.19) .. (307.72,224.99) .. controls (307.72,225.79) and (306.96,226.44) .. (306.03,226.44) .. controls (305.1,226.44) and (304.34,225.79) .. (304.34,224.99) -- cycle ;
%Shape: Ellipse [id:dp9592402529849944]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (317.2,191.73) .. controls (317.2,190.93) and (317.95,190.28) .. (318.89,190.28) .. controls (319.82,190.28) and (320.57,190.93) .. (320.57,191.73) .. controls (320.57,192.53) and (319.82,193.18) .. (318.89,193.18) .. controls (317.95,193.18) and (317.2,192.53) .. (317.2,191.73) -- cycle ;
%Shape: Ellipse [id:dp21535542685573428]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (317.89,197.24) .. controls (317.89,196.44) and (318.65,195.79) .. (319.58,195.79) .. controls (320.52,195.79) and (321.27,196.44) .. (321.27,197.24) .. controls (321.27,198.04) and (320.52,198.69) .. (319.58,198.69) .. controls (318.65,198.69) and (317.89,198.04) .. (317.89,197.24) -- cycle ;
%Shape: Ellipse [id:dp43454667067156816]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (318.17,203.47) .. controls (318.17,202.67) and (318.93,202.02) .. (319.86,202.02) .. controls (320.79,202.02) and (321.55,202.67) .. (321.55,203.47) .. controls (321.55,204.27) and (320.79,204.92) .. (319.86,204.92) .. controls (318.93,204.92) and (318.17,204.27) .. (318.17,203.47) -- cycle ;
%Shape: Ellipse [id:dp9701948912141243]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (318.93,210.24) .. controls (318.93,209.44) and (319.68,208.79) .. (320.62,208.79) .. controls (321.55,208.79) and (322.3,209.44) .. (322.3,210.24) .. controls (322.3,211.04) and (321.55,211.69) .. (320.62,211.69) .. controls (319.68,211.69) and (318.93,211.04) .. (318.93,210.24) -- cycle ;
%Shape: Ellipse [id:dp5208741511315487]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (319.29,217.12) .. controls (319.29,216.32) and (320.05,215.67) .. (320.98,215.67) .. controls (321.91,215.67) and (322.67,216.32) .. (322.67,217.12) .. controls (322.67,217.92) and (321.91,218.57) .. (320.98,218.57) .. controls (320.05,218.57) and (319.29,217.92) .. (319.29,217.12) -- cycle ;
%Shape: Ellipse [id:dp9960070270987895]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (320.13,224.91) .. controls (320.13,224.11) and (320.88,223.46) .. (321.82,223.46) .. controls (322.75,223.46) and (323.5,224.11) .. (323.5,224.91) .. controls (323.5,225.71) and (322.75,226.36) .. (321.82,226.36) .. controls (320.88,226.36) and (320.13,225.71) .. (320.13,224.91) -- cycle ;
%Shape: Ellipse [id:dp08258824726895897]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (331.87,192.13) .. controls (331.87,191.33) and (332.62,190.68) .. (333.56,190.68) .. controls (334.49,190.68) and (335.25,191.33) .. (335.25,192.13) .. controls (335.25,192.93) and (334.49,193.58) .. (333.56,193.58) .. controls (332.62,193.58) and (331.87,192.93) .. (331.87,192.13) -- cycle ;
%Shape: Ellipse [id:dp5433546755374361]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (333.51,197.36) .. controls (333.51,196.56) and (334.26,195.91) .. (335.19,195.91) .. controls (336.13,195.91) and (336.88,196.56) .. (336.88,197.36) .. controls (336.88,198.16) and (336.13,198.81) .. (335.19,198.81) .. controls (334.26,198.81) and (333.51,198.16) .. (333.51,197.36) -- cycle ;
%Shape: Ellipse [id:dp7579405737158253]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (334.36,203.23) .. controls (334.36,202.43) and (335.12,201.78) .. (336.05,201.78) .. controls (336.98,201.78) and (337.74,202.43) .. (337.74,203.23) .. controls (337.74,204.03) and (336.98,204.68) .. (336.05,204.68) .. controls (335.12,204.68) and (334.36,204.03) .. (334.36,203.23) -- cycle ;
%Shape: Ellipse [id:dp9171787071340098]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (335.51,210.5) .. controls (335.51,209.7) and (336.26,209.05) .. (337.19,209.05) .. controls (338.13,209.05) and (338.88,209.7) .. (338.88,210.5) .. controls (338.88,211.3) and (338.13,211.95) .. (337.19,211.95) .. controls (336.26,211.95) and (335.51,211.3) .. (335.51,210.5) -- cycle ;
%Shape: Ellipse [id:dp5333550722820979]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (336.32,217.12) .. controls (336.32,216.32) and (337.07,215.67) .. (338,215.67) .. controls (338.94,215.67) and (339.69,216.32) .. (339.69,217.12) .. controls (339.69,217.92) and (338.94,218.57) .. (338,218.57) .. controls (337.07,218.57) and (336.32,217.92) .. (336.32,217.12) -- cycle ;
%Shape: Ellipse [id:dp9796837086495329]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (337.99,225.03) .. controls (337.99,224.23) and (338.75,223.58) .. (339.68,223.58) .. controls (340.61,223.58) and (341.37,224.23) .. (341.37,225.03) .. controls (341.37,225.83) and (340.61,226.48) .. (339.68,226.48) .. controls (338.75,226.48) and (337.99,225.83) .. (337.99,225.03) -- cycle ;
%Shape: Ellipse [id:dp7978787114926011]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (346.92,191.49) .. controls (346.92,190.69) and (347.68,190.04) .. (348.61,190.04) .. controls (349.54,190.04) and (350.3,190.69) .. (350.3,191.49) .. controls (350.3,192.29) and (349.54,192.94) .. (348.61,192.94) .. controls (347.68,192.94) and (346.92,192.29) .. (346.92,191.49) -- cycle ;
%Shape: Ellipse [id:dp8545275834894772]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (348.6,197.24) .. controls (348.6,196.44) and (349.35,195.79) .. (350.28,195.79) .. controls (351.22,195.79) and (351.97,196.44) .. (351.97,197.24) .. controls (351.97,198.04) and (351.22,198.69) .. (350.28,198.69) .. controls (349.35,198.69) and (348.6,198.04) .. (348.6,197.24) -- cycle ;
%Shape: Ellipse [id:dp9024642638477771]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (349.99,203.35) .. controls (349.99,202.55) and (350.75,201.9) .. (351.68,201.9) .. controls (352.61,201.9) and (353.37,202.55) .. (353.37,203.35) .. controls (353.37,204.15) and (352.61,204.8) .. (351.68,204.8) .. controls (350.75,204.8) and (349.99,204.15) .. (349.99,203.35) -- cycle ;
%Shape: Ellipse [id:dp139500631439212]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (351.93,210.1) .. controls (351.93,209.3) and (352.69,208.65) .. (353.62,208.65) .. controls (354.55,208.65) and (355.31,209.3) .. (355.31,210.1) .. controls (355.31,210.9) and (354.55,211.55) .. (353.62,211.55) .. controls (352.69,211.55) and (351.93,210.9) .. (351.93,210.1) -- cycle ;
%Shape: Ellipse [id:dp053214074310489634]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (353.76,217.24) .. controls (353.76,216.44) and (354.52,215.79) .. (355.45,215.79) .. controls (356.38,215.79) and (357.14,216.44) .. (357.14,217.24) .. controls (357.14,218.04) and (356.38,218.69) .. (355.45,218.69) .. controls (354.52,218.69) and (353.76,218.04) .. (353.76,217.24) -- cycle ;
%Shape: Ellipse [id:dp40952769247810483]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (355.71,225.03) .. controls (355.71,224.23) and (356.47,223.58) .. (357.4,223.58) .. controls (358.33,223.58) and (359.09,224.23) .. (359.09,225.03) .. controls (359.09,225.83) and (358.33,226.48) .. (357.4,226.48) .. controls (356.47,226.48) and (355.71,225.83) .. (355.71,225.03) -- cycle ;
%Shape: Ellipse [id:dp6883505384870545]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (360.46,191.49) .. controls (360.46,190.69) and (361.21,190.04) .. (362.15,190.04) .. controls (363.08,190.04) and (363.84,190.69) .. (363.84,191.49) .. controls (363.84,192.29) and (363.08,192.94) .. (362.15,192.94) .. controls (361.21,192.94) and (360.46,192.29) .. (360.46,191.49) -- cycle ;
%Shape: Ellipse [id:dp8228987503250764]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (362.69,197) .. controls (362.69,196.2) and (363.45,195.55) .. (364.38,195.55) .. controls (365.31,195.55) and (366.07,196.2) .. (366.07,197) .. controls (366.07,197.8) and (365.31,198.45) .. (364.38,198.45) .. controls (363.45,198.45) and (362.69,197.8) .. (362.69,197) -- cycle ;
%Shape: Ellipse [id:dp18713358672003966]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (364.92,203.35) .. controls (364.92,202.55) and (365.68,201.9) .. (366.61,201.9) .. controls (367.54,201.9) and (368.3,202.55) .. (368.3,203.35) .. controls (368.3,204.15) and (367.54,204.8) .. (366.61,204.8) .. controls (365.68,204.8) and (364.92,204.15) .. (364.92,203.35) -- cycle ;
%Shape: Ellipse [id:dp879320396266565]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (367.44,210.5) .. controls (367.44,209.7) and (368.19,209.05) .. (369.12,209.05) .. controls (370.06,209.05) and (370.81,209.7) .. (370.81,210.5) .. controls (370.81,211.3) and (370.06,211.95) .. (369.12,211.95) .. controls (368.19,211.95) and (367.44,211.3) .. (367.44,210.5) -- cycle ;
%Shape: Ellipse [id:dp8350403435643714]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (369.53,217) .. controls (369.53,216.2) and (370.28,215.55) .. (371.22,215.55) .. controls (372.15,215.55) and (372.91,216.2) .. (372.91,217) .. controls (372.91,217.8) and (372.15,218.45) .. (371.22,218.45) .. controls (370.28,218.45) and (369.53,217.8) .. (369.53,217) -- cycle ;
%Shape: Ellipse [id:dp17300285054057252]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (372.46,224.67) .. controls (372.46,223.87) and (373.22,223.22) .. (374.15,223.22) .. controls (375.08,223.22) and (375.84,223.87) .. (375.84,224.67) .. controls (375.84,225.47) and (375.08,226.12) .. (374.15,226.12) .. controls (373.22,226.12) and (372.46,225.47) .. (372.46,224.67) -- cycle ;
%Shape: Ellipse [id:dp6550411424766795]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (374.97,191.49) .. controls (374.97,190.69) and (375.73,190.04) .. (376.66,190.04) .. controls (377.59,190.04) and (378.35,190.69) .. (378.35,191.49) .. controls (378.35,192.29) and (377.59,192.94) .. (376.66,192.94) .. controls (375.73,192.94) and (374.97,192.29) .. (374.97,191.49) -- cycle ;
%Shape: Ellipse [id:dp2665045036715892]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (377.48,197.12) .. controls (377.48,196.32) and (378.24,195.67) .. (379.17,195.67) .. controls (380.1,195.67) and (380.86,196.32) .. (380.86,197.12) .. controls (380.86,197.92) and (380.1,198.57) .. (379.17,198.57) .. controls (378.24,198.57) and (377.48,197.92) .. (377.48,197.12) -- cycle ;
%Shape: Ellipse [id:dp211226681996602]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (379.86,203.23) .. controls (379.86,202.43) and (380.61,201.78) .. (381.54,201.78) .. controls (382.48,201.78) and (383.23,202.43) .. (383.23,203.23) .. controls (383.23,204.03) and (382.48,204.68) .. (381.54,204.68) .. controls (380.61,204.68) and (379.86,204.03) .. (379.86,203.23) -- cycle ;
%Shape: Ellipse [id:dp5135394266621403]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (382.65,209.82) .. controls (382.65,209.02) and (383.4,208.37) .. (384.34,208.37) .. controls (385.27,208.37) and (386.02,209.02) .. (386.02,209.82) .. controls (386.02,210.62) and (385.27,211.27) .. (384.34,211.27) .. controls (383.4,211.27) and (382.65,210.62) .. (382.65,209.82) -- cycle ;
%Shape: Ellipse [id:dp4800412429140355]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (386,217) .. controls (386,216.2) and (386.75,215.55) .. (387.68,215.55) .. controls (388.62,215.55) and (389.37,216.2) .. (389.37,217) .. controls (389.37,217.8) and (388.62,218.45) .. (387.68,218.45) .. controls (386.75,218.45) and (386,217.8) .. (386,217) -- cycle ;
%Shape: Ellipse [id:dp9897837661968554]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (389.21,224.67) .. controls (389.21,223.87) and (389.96,223.22) .. (390.89,223.22) .. controls (391.83,223.22) and (392.58,223.87) .. (392.58,224.67) .. controls (392.58,225.47) and (391.83,226.12) .. (390.89,226.12) .. controls (389.96,226.12) and (389.21,225.47) .. (389.21,224.67) -- cycle ;
%Shape: Ellipse [id:dp9164081127827948]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (387.39,191.49) .. controls (387.39,190.69) and (388.15,190.04) .. (389.08,190.04) .. controls (390.01,190.04) and (390.77,190.69) .. (390.77,191.49) .. controls (390.77,192.29) and (390.01,192.94) .. (389.08,192.94) .. controls (388.15,192.94) and (387.39,192.29) .. (387.39,191.49) -- cycle ;
%Shape: Ellipse [id:dp5397824783956662]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (390.18,197.12) .. controls (390.18,196.32) and (390.94,195.67) .. (391.87,195.67) .. controls (392.8,195.67) and (393.56,196.32) .. (393.56,197.12) .. controls (393.56,197.92) and (392.8,198.57) .. (391.87,198.57) .. controls (390.94,198.57) and (390.18,197.92) .. (390.18,197.12) -- cycle ;
%Shape: Ellipse [id:dp09820896087807829]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (393.53,203.35) .. controls (393.53,202.55) and (394.29,201.9) .. (395.22,201.9) .. controls (396.15,201.9) and (396.91,202.55) .. (396.91,203.35) .. controls (396.91,204.15) and (396.15,204.8) .. (395.22,204.8) .. controls (394.29,204.8) and (393.53,204.15) .. (393.53,203.35) -- cycle ;
%Shape: Ellipse [id:dp9648161409230518]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (397.52,210.32) .. controls (397.52,209.52) and (398.28,208.87) .. (399.21,208.87) .. controls (400.14,208.87) and (400.9,209.52) .. (400.9,210.32) .. controls (400.9,211.12) and (400.14,211.77) .. (399.21,211.77) .. controls (398.28,211.77) and (397.52,211.12) .. (397.52,210.32) -- cycle ;
%Shape: Ellipse [id:dp9189826271468595]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (401.35,216.88) .. controls (401.35,216.08) and (402.1,215.43) .. (403.04,215.43) .. controls (403.97,215.43) and (404.72,216.08) .. (404.72,216.88) .. controls (404.72,217.68) and (403.97,218.33) .. (403.04,218.33) .. controls (402.1,218.33) and (401.35,217.68) .. (401.35,216.88) -- cycle ;
%Shape: Ellipse [id:dp356684240956189]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (405.67,224.67) .. controls (405.67,223.87) and (406.43,223.22) .. (407.36,223.22) .. controls (408.29,223.22) and (409.05,223.87) .. (409.05,224.67) .. controls (409.05,225.47) and (408.29,226.12) .. (407.36,226.12) .. controls (406.43,226.12) and (405.67,225.47) .. (405.67,224.67) -- cycle ;
%Shape: Ellipse [id:dp30965018239062236]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (401.21,191.49) .. controls (401.21,190.69) and (401.96,190.04) .. (402.9,190.04) .. controls (403.83,190.04) and (404.58,190.69) .. (404.58,191.49) .. controls (404.58,192.29) and (403.83,192.94) .. (402.9,192.94) .. controls (401.96,192.94) and (401.21,192.29) .. (401.21,191.49) -- cycle ;
%Shape: Ellipse [id:dp268329300327677]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (404,197) .. controls (404,196.2) and (404.75,195.55) .. (405.69,195.55) .. controls (406.62,195.55) and (407.38,196.2) .. (407.38,197) .. controls (407.38,197.8) and (406.62,198.45) .. (405.69,198.45) .. controls (404.75,198.45) and (404,197.8) .. (404,197) -- cycle ;
%Shape: Ellipse [id:dp21056106596973323]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (407.63,203.23) .. controls (407.63,202.43) and (408.38,201.78) .. (409.32,201.78) .. controls (410.25,201.78) and (411,202.43) .. (411,203.23) .. controls (411,204.03) and (410.25,204.68) .. (409.32,204.68) .. controls (408.38,204.68) and (407.63,204.03) .. (407.63,203.23) -- cycle ;
%Shape: Ellipse [id:dp9269572031249087]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (411.49,210.24) .. controls (411.49,209.44) and (412.25,208.79) .. (413.18,208.79) .. controls (414.11,208.79) and (414.87,209.44) .. (414.87,210.24) .. controls (414.87,211.04) and (414.11,211.69) .. (413.18,211.69) .. controls (412.25,211.69) and (411.49,211.04) .. (411.49,210.24) -- cycle ;
%Shape: Ellipse [id:dp816663150550079]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (415.16,217) .. controls (415.16,216.2) and (415.92,215.55) .. (416.85,215.55) .. controls (417.78,215.55) and (418.54,216.2) .. (418.54,217) .. controls (418.54,217.8) and (417.78,218.45) .. (416.85,218.45) .. controls (415.92,218.45) and (415.16,217.8) .. (415.16,217) -- cycle ;
%Shape: Ellipse [id:dp7731814355118924]
\draw  [fill={rgb, 255:red, 118; green, 113; blue, 113 }  ,fill opacity=1 ] (419.35,224.79) .. controls (419.35,223.99) and (420.11,223.34) .. (421.04,223.34) .. controls (421.97,223.34) and (422.73,223.99) .. (422.73,224.79) .. controls (422.73,225.59) and (421.97,226.24) .. (421.04,226.24) .. controls (420.11,226.24) and (419.35,225.59) .. (419.35,224.79) -- cycle ;
%Straight Lines [id:da3000520596553806]
\draw    (196.12,186.91) -- (168.03,233.62) ;
%Straight Lines [id:da163813348699974]
\draw    (168.03,233.62) -- (171.25,233.56) ;
%Straight Lines [id:da5727664193245703]
\draw    (196.12,186.91) -- (199.34,186.85) ;
%Straight Lines [id:da6793044911060069]
\draw    (440.92,233.12) -- (444.14,233.05) ;
%Straight Lines [id:da7452026016611717]
\draw    (413.8,187) -- (417.02,186.94) ;
%Straight Lines [id:da7007647307003126]
\draw    (417.02,186.94) -- (444.14,233.05) ;
%Straight Lines [id:da9237803080397062]
\draw [color={rgb, 255:red, 162; green, 157; blue, 157 }  ,draw opacity=1 ]   (179.76,177.9) -- (199.23,177.7) ;
\draw [shift={(202.23,177.66)}, rotate = 539.39] [fill={rgb, 255:red, 162; green, 157; blue, 157 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (6.25,-3) -- (0,0) -- (6.25,3) -- cycle    ;
%Straight Lines [id:da8131104182423468]
\draw [color={rgb, 255:red, 162; green, 157; blue, 157 }  ,draw opacity=1 ]   (179.76,177.9) -- (170.65,188.89) ;
\draw [shift={(168.74,191.2)}, rotate = 309.65999999999997] [fill={rgb, 255:red, 162; green, 157; blue, 157 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (6.25,-3) -- (0,0) -- (6.25,3) -- cycle    ;
%Shape: Wave [id:dp5678827377050879]
\draw  [color={rgb, 255:red, 229; green, 21; blue, 21 }  ,draw opacity=1 ][line width=1.5]  (309.6,198.6) .. controls (324.2,194.36) and (338.1,190.3) .. (338.1,185.6) .. controls (338.1,180.9) and (324.2,176.84) .. (309.6,172.6) .. controls (295,168.36) and (281.1,164.3) .. (281.1,159.6) .. controls (281.1,154.9) and (295,150.84) .. (309.6,146.6) .. controls (324.2,142.36) and (338.1,138.3) .. (338.1,133.6) .. controls (338.1,128.9) and (324.2,124.84) .. (309.6,120.6) .. controls (295,116.36) and (281.1,112.3) .. (281.1,107.6) .. controls (281.1,102.95) and (294.67,98.94) .. (309.08,94.75) ;
%Straight Lines [id:da29341364388127156]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (309.8,100.6) -- (297.2,100.6) ;
\draw [shift={(294.2,100.6)}, rotate = 360] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da5859632848227461]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (310,114.75) -- (297.2,114.63) ;
\draw [shift={(294.2,114.6)}, rotate = 360.53999999999996] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da15254151544652794]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (310.6,107.2) -- (285,107.24) ;
\draw [shift={(282,107.25)}, rotate = 359.9] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da97623530444903]
\draw [color={rgb, 255:red, 162; green, 157; blue, 157 }  ,draw opacity=1 ]   (179.76,177.9) -- (179.8,158) ;
\draw [shift={(179.8,155)}, rotate = 450.09] [fill={rgb, 255:red, 162; green, 157; blue, 157 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (6.25,-3) -- (0,0) -- (6.25,3) -- cycle    ;
%Straight Lines [id:da3776683501360716]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (310.2,140.6) -- (321.6,140.6) ;
\draw [shift={(324.6,140.6)}, rotate = 180] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da22435064621656453]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (310.6,126.6) -- (323.6,126.6) ;
\draw [shift={(326.6,126.6)}, rotate = 180] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da4117346036483176]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (309.8,134) -- (334.5,133.78) ;
\draw [shift={(337.5,133.75)}, rotate = 539.48] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da4639690103049321]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (309.8,152.2) -- (297.2,152.2) ;
\draw [shift={(294.2,152.2)}, rotate = 360] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da9776464579169482]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (310.2,165.8) -- (297.2,166.13) ;
\draw [shift={(294.2,166.2)}, rotate = 358.57] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da2751823416687842]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (310.6,158.8) -- (284.4,159.34) ;
\draw [shift={(281.4,159.4)}, rotate = 358.82] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da5393896354953414]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (309,192.6) -- (320.4,192.6) ;
\draw [shift={(323.4,192.6)}, rotate = 180] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da015970204232534035]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (309.4,178.6) -- (322.4,178.6) ;
\draw [shift={(325.4,178.6)}, rotate = 180] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da6560045675946828]
\draw [color={rgb, 255:red, 227; green, 180; blue, 187 }  ,draw opacity=1 ]   (309.6,186) -- (333.8,185.47) ;
\draw [shift={(336.8,185.4)}, rotate = 538.74] [fill={rgb, 255:red, 227; green, 180; blue, 187 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (3.57,-1.72) -- (0,0) -- (3.57,1.72) -- cycle    ;
%Straight Lines [id:da6731118036698975]
\draw [color={rgb, 255:red, 20; green, 88; blue, 145 }  ,draw opacity=1 ][line width=1.5]    (249.92,200.08) -- (249.78,154.23) ;
\draw [shift={(249.77,150.23)}, rotate = 449.82] [fill={rgb, 255:red, 20; green, 88; blue, 145 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (9.29,-4.46) -- (0,0) -- (9.29,4.46) -- cycle    ;
%Straight Lines [id:da47217219479835637]
\draw [color={rgb, 255:red, 20; green, 88; blue, 145 }  ,draw opacity=1 ][line width=1.5]    (270.23,200.08) -- (270.23,154.23) ;
\draw [shift={(270.23,150.23)}, rotate = 450] [fill={rgb, 255:red, 20; green, 88; blue, 145 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (9.29,-4.46) -- (0,0) -- (9.29,4.46) -- cycle    ;
%Straight Lines [id:da5501938249838205]
\draw [color={rgb, 255:red, 20; green, 88; blue, 145 }  ,draw opacity=1 ][line width=1.5]    (310.38,199.92) -- (310.24,154.23) ;
\draw [shift={(310.23,150.23)}, rotate = 449.82] [fill={rgb, 255:red, 20; green, 88; blue, 145 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (9.29,-4.46) -- (0,0) -- (9.29,4.46) -- cycle    ;
%Straight Lines [id:da0658558878821176]
\draw [color={rgb, 255:red, 20; green, 88; blue, 145 }  ,draw opacity=1 ][line width=1.5]    (330.23,200.08) -- (330.09,154.38) ;
\draw [shift={(330.08,150.38)}, rotate = 449.82] [fill={rgb, 255:red, 20; green, 88; blue, 145 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (9.29,-4.46) -- (0,0) -- (9.29,4.46) -- cycle    ;
%Straight Lines [id:da9765507457725142]
\draw [color={rgb, 255:red, 20; green, 88; blue, 145 }  ,draw opacity=1 ][line width=1.5]    (350.08,200.23) -- (349.94,154.54) ;
\draw [shift={(349.92,150.54)}, rotate = 449.82] [fill={rgb, 255:red, 20; green, 88; blue, 145 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (9.29,-4.46) -- (0,0) -- (9.29,4.46) -- cycle    ;
%Straight Lines [id:da6498342095486176]
\draw [color={rgb, 255:red, 20; green, 88; blue, 145 }  ,draw opacity=1 ][line width=1.5]    (290.38,200.08) -- (290.24,154.38) ;
\draw [shift={(290.23,150.38)}, rotate = 449.82] [fill={rgb, 255:red, 20; green, 88; blue, 145 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (9.29,-4.46) -- (0,0) -- (9.29,4.46) -- cycle    ;

% Text Node
\draw (147.56,83.2) node [anchor=north west][inner sep=0.75pt]  [font=\normalsize] [align=left] {\begin{minipage}[lt]{84.35pt}\setlength\topsep{0pt}
\begin{center}
\textcolor[rgb]{0.82,0.01,0.11}{{\fontfamily{helvet}\selectfont Elctric Field of the}}\\\textcolor[rgb]{0.82,0.01,0.11}{{\fontfamily{helvet}\selectfont Dressing Field (\textbf{E})}}
\end{center}

\end{minipage}};
% Text Node
\draw (358.46,156.94) node [anchor=north west][inner sep=0.75pt]  [font=\normalsize] [align=left] {\begin{minipage}[lt]{86.06pt}\setlength\topsep{0pt}
\begin{center}
\textcolor[rgb]{0.08,0.35,0.57}{{\fontfamily{helvet}\selectfont Magnetic Field (\textbf{B})}}
\end{center}

\end{minipage}};
% Text Node
\draw (168,134.9) node [anchor=north west][inner sep=0.75pt]    {$z$};
% Text Node
\draw (154,175.9) node [anchor=north west][inner sep=0.75pt]    {$x$};
% Text Node
\draw (204.5,156.9) node [anchor=north west][inner sep=0.75pt]    {$y$};
% Text Node
\draw (441,208) node [anchor=north west][inner sep=0.75pt]   [align=left] {2DEG};

\end{tikzpicture}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%COPY HERE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}