### 2021-07-01 paper/v1.0/sections/4_floquet_fermi_golden.tex
Corrections

In this section we are going to derive the Floquet Fermi goldern rule for previously derived quantum Floquet states using $t-t'$ formalism \cite{bruus04}.