### 2021-07-12 paper/v1.0/abstract.tex
Change abstract paper

A generalized mathematical model for the transport properties of systems exposed to a stationary magnetic and a strong electromagnetic field is presented. The new formulation, which applies to the two-dimensional dressed quantum Hall systems, is based on Landau quantization theory and Floquet-Drude conductivity method. We model our system as a two-dimensional electron gas (2DEG) that interacts with two external fields. To incorporate the strong light coupling with the 2DEG, we utilize the Floquet theory to analyze the effect non perturbatively. Moreover, the Floquet Fermi "golden rule" is adopted to explore the scattering effects for Floquet states in disordered quantum Hall systems. Based on our fully analytical expression and particular graphical representations, we demonstrate
that the characteristics of conductivities in two-dimensional quantum Hall systems can manipulate using a dressed field. The outcomes align with theoretical descriptions which are already well-suited with experimental results at the same time our theory provides a more generalized analysis on the properties of conductivity in quantum Hall systems. Thus, this model more realistically describes that how to use external strong radiation as a tool to utilize transport properties in various 2D nanostructures which serve as a basis for nano-optoelectronic devices.