### 2021-07-18 theory/figures/fig_2new.tex

\documentclass{standalone}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tikz,pgf} %and any other packages or tikzlibraries your picture needs
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{physics}
\usepackage{amssymb}
\usepackage{mathtools}
\usetikzlibrary{shapes,arrows,patterns}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%COPY HERE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Pattern Info

\tikzset{
pattern size/.store in=\mcSize,
pattern size = 5pt,
pattern thickness/.store in=\mcThickness,
pattern thickness = 0.3pt,
pattern radius/.store in=\mcRadius,
pattern radius = 1pt}
\makeatletter
\pgfutil@ifundefined{pgf@pattern@name@_sbtxotmvc}{
\pgfdeclarepatternformonly[\mcThickness,\mcSize]{_sbtxotmvc}
{\pgfqpoint{0pt}{-\mcThickness}}
{\pgfpoint{\mcSize}{\mcSize}}
{\pgfpoint{\mcSize}{\mcSize}}
{
\pgfsetcolor{\tikz@pattern@color}
\pgfsetlinewidth{\mcThickness}
\pgfpathmoveto{\pgfqpoint{0pt}{\mcSize}}
\pgfpathlineto{\pgfpoint{\mcSize+\mcThickness}{-\mcThickness}}
\pgfusepath{stroke}
}}
\makeatother

% Pattern Info

\tikzset{
pattern size/.store in=\mcSize,
pattern size = 5pt,
pattern thickness/.store in=\mcThickness,
pattern thickness = 0.3pt,
pattern radius/.store in=\mcRadius,
pattern radius = 1pt}
\makeatletter
\pgfutil@ifundefined{pgf@pattern@name@_6w8gp62zx}{
\pgfdeclarepatternformonly[\mcThickness,\mcSize]{_6w8gp62zx}
{\pgfqpoint{0pt}{0pt}}
{\pgfpoint{\mcSize+\mcThickness}{\mcSize+\mcThickness}}
{\pgfpoint{\mcSize}{\mcSize}}
{
\pgfsetcolor{\tikz@pattern@color}
\pgfsetlinewidth{\mcThickness}
\pgfpathmoveto{\pgfqpoint{0pt}{0pt}}
\pgfpathlineto{\pgfpoint{\mcSize+\mcThickness}{\mcSize+\mcThickness}}
\pgfusepath{stroke}
}}
\makeatother

% Pattern Info

\tikzset{
pattern size/.store in=\mcSize,
pattern size = 5pt,
pattern thickness/.store in=\mcThickness,
pattern thickness = 0.3pt,
pattern radius/.store in=\mcRadius,
pattern radius = 1pt}
\makeatletter
\pgfutil@ifundefined{pgf@pattern@name@_zm822ewj2}{
\pgfdeclarepatternformonly[\mcThickness,\mcSize]{_zm822ewj2}
{\pgfqpoint{0pt}{0pt}}
{\pgfpoint{\mcSize+\mcThickness}{\mcSize+\mcThickness}}
{\pgfpoint{\mcSize}{\mcSize}}
{
\pgfsetcolor{\tikz@pattern@color}
\pgfsetlinewidth{\mcThickness}
\pgfpathmoveto{\pgfqpoint{0pt}{0pt}}
\pgfpathlineto{\pgfpoint{\mcSize+\mcThickness}{\mcSize+\mcThickness}}
\pgfusepath{stroke}
}}
\makeatother
\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300

%Shape: Rectangle [id:dp9138712933038846]
\draw  [color={rgb, 255:red, 255; green, 255; blue, 255 }  ,draw opacity=1 ][pattern=_sbtxotmvc,pattern size=5.550000000000001pt,pattern thickness=0.75pt,pattern radius=0pt, pattern color={rgb, 255:red, 230; green, 228; blue, 228}] (210.57,130.37) -- (429.86,130.37) -- (429.86,169.4) -- (210.57,169.4) -- cycle ;
%Straight Lines [id:da062381655627882227]
\draw [line width=0.75]    (320.14,216.43) -- (319.86,93.43) ;
\draw [shift={(319.86,90.43)}, rotate = 449.87] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (5.36,-2.57) -- (0,0) -- (5.36,2.57) -- cycle    ;
%Shape: Boxed Line [id:dp7700428726953505]
\draw [line width=0.75]    (210.14,210.14) -- (443.6,210.42) ;
\draw [shift={(446.6,210.43)}, rotate = 180.07] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (5.36,-2.57) -- (0,0) -- (5.36,2.57) -- cycle    ;
%Straight Lines [id:da1183130111686681]
\draw [color={rgb, 255:red, 15; green, 77; blue, 150 }  ,draw opacity=1 ]   (210.26,190.14) -- (429.97,189.86) ;
%Straight Lines [id:da9178920527162777]
\draw [color={rgb, 255:red, 15; green, 77; blue, 150 }  ,draw opacity=1 ]   (210.66,150.14) -- (430.66,150.14) ;
%Straight Lines [id:da18288396037355947]
\draw [color={rgb, 255:red, 15; green, 77; blue, 150 }  ,draw opacity=1 ]   (209.46,110.54) -- (429.46,110.54) ;
%Straight Lines [id:da7423332306251238]
\draw [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][line width=0.75]  [dash pattern={on 4.5pt off 4.5pt}]  (210.29,129.8) -- (429.86,130.37) ;
%Straight Lines [id:da5380911943955708]
\draw [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ] [dash pattern={on 4.5pt off 4.5pt}]  (216,169.11) -- (435.57,169.69) ;
%Straight Lines [id:da5108918794430821]
\draw [color={rgb, 255:red, 245; green, 166; blue, 35 }  ,draw opacity=1 ]   (210.29,139.94) -- (430.14,140.37) ;
%Shape: Circle [id:dp6526443567602316]
\draw  [pattern=_6w8gp62zx,pattern size=1.9500000000000002pt,pattern thickness=0.75pt,pattern radius=0pt, pattern color={rgb, 255:red, 155; green, 155; blue, 155}] (346.39,148.12) .. controls (347.41,146.29) and (349.72,145.64) .. (351.55,146.66) .. controls (353.37,147.68) and (354.02,149.99) .. (353,151.81) .. controls (351.98,153.64) and (349.67,154.29) .. (347.85,153.27) .. controls (346.02,152.25) and (345.37,149.94) .. (346.39,148.12) -- cycle ;
%Shape: Circle [id:dp3302958539355392]
\draw  [pattern=_zm822ewj2,pattern size=1.9500000000000002pt,pattern thickness=0.75pt,pattern radius=0pt, pattern color={rgb, 255:red, 155; green, 155; blue, 155}] (397.89,140.56) .. controls (397.89,138.47) and (399.58,136.77) .. (401.67,136.77) .. controls (403.76,136.77) and (405.46,138.47) .. (405.46,140.56) .. controls (405.46,142.65) and (403.76,144.34) .. (401.67,144.34) .. controls (399.58,144.34) and (397.89,142.65) .. (397.89,140.56) -- cycle ;
%Curve Lines [id:da4768841907072914]
\draw [color={rgb, 255:red, 200; green, 31; blue, 51 }  ,draw opacity=1 ]   (349.86,145) .. controls (353.28,116.82) and (383.19,108.39) .. (397.87,129.76) ;
\draw [shift={(399.4,132.2)}, rotate = 240.15] [fill={rgb, 255:red, 200; green, 31; blue, 51 }  ,fill opacity=1 ][line width=0.08]  [draw opacity=0] (7.14,-3.43) -- (0,0) -- (7.14,3.43) -- cycle    ;
%Straight Lines [id:da804813251167583]
\draw    (350.14,205) -- (350.29,215.86) ;
%Straight Lines [id:da9646977497816096]
\draw    (379.92,205.11) -- (380.06,215.97) ;
%Straight Lines [id:da10347201047155785]
\draw    (360.14,205) -- (360.29,215.86) ;
%Straight Lines [id:da030579557349847963]
\draw    (370.14,205.57) -- (370.29,216.43) ;
%Straight Lines [id:da5754192540908345]
\draw    (390.14,204.56) -- (390.29,215.41) ;
%Straight Lines [id:da7881344991506445]
\draw    (419.92,204.67) -- (420.06,215.52) ;
%Straight Lines [id:da5081549699459584]
\draw    (400.14,204.56) -- (400.29,215.41) ;
%Straight Lines [id:da7295740773263839]
\draw    (410.14,205.13) -- (410.29,215.98) ;
%Straight Lines [id:da29498141413358026]
\draw    (270.14,205.22) -- (270.29,216.08) ;
%Straight Lines [id:da05159592463246754]
\draw    (299.92,205.33) -- (300.06,216.19) ;
%Straight Lines [id:da8075283207227479]
\draw    (280.14,205.22) -- (280.29,216.08) ;
%Straight Lines [id:da5383953202315974]
\draw    (290.14,205.79) -- (290.29,216.65) ;
%Straight Lines [id:da4613327018299631]
\draw    (310.14,204.78) -- (310.29,215.63) ;
%Straight Lines [id:da2686333335167921]
\draw    (339.92,204.89) -- (340.06,215.75) ;
%Straight Lines [id:da00021650773652504007]
\draw    (330.14,205.35) -- (330.29,216.21) ;
%Straight Lines [id:da02622977015161565]
\draw    (219.48,204.67) -- (219.62,215.52) ;
%Straight Lines [id:da36110488827684883]
\draw    (229.7,204.11) -- (229.84,214.97) ;
%Straight Lines [id:da8472813973801454]
\draw    (259.48,204.22) -- (259.62,215.08) ;
%Straight Lines [id:da674226541094751]
\draw    (239.7,204.11) -- (239.84,214.97) ;
%Straight Lines [id:da7672438249417022]
\draw    (249.7,204.68) -- (249.84,215.54) ;
%Straight Lines [id:da0871509351896862]
\draw    (429.92,204.82) -- (430.06,215.68) ;

% Text Node
\draw (319.71,74.34) node [anchor=north west][inner sep=0.75pt]    {$\text{Energy}$};
% Text Node
\draw (449.71,201.29) node [anchor=north west][inner sep=0.75pt]    {$k_{x}$};
% Text Node
\draw (434.66,145.77) node [anchor=north west][inner sep=0.75pt]  [color={rgb, 255:red, 15; green, 77; blue, 150 }  ,opacity=1 ]  {$\varepsilon _{\beta }$};
% Text Node
\draw (196.29,134.31) node [anchor=north west][inner sep=0.75pt]  [color={rgb, 255:red, 245; green, 166; blue, 35 }  ,opacity=1 ]  {$\varepsilon $};
% Text Node
\draw (335.2,152.77) node [anchor=north west][inner sep=0.75pt]  [font=\normalsize]  {$\ket{\psi _{\beta }}$};
% Text Node
\draw (397.49,113.6) node [anchor=north west][inner sep=0.75pt]  [font=\normalsize]  {$\ket{\Psi _{\alpha }}$};

\end{tikzpicture}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%COPY HERE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}