### 2021-08-02 paper/v1.0/abstract.tex
Formatting and correcting

A generalized mathematical model for predicting the transport properties of a quantum system exposed to a stationary magnetic field and a high intensity electromagnetic field is presented.
The new formulation, which applies to two-dimensional(2D) dressed quantum Hall systems, is based on Landau quantization theory and Floquet-Drude conductivity approach.
We model our system as a two-dimensional electron gas (2DEG) that interacts with two external fields. To analyze the strong light coupling with the 2DEG, we employ the Floquet theory as a nonperturbative procedure.
Moreover, the Floquet Fermi golden rule is adopted to explore the impurity scattering effects on charge transport in disordered quantum Hall systems.
We derive fully analytical expressions to describe longitudinal components in the conductivity tensor in dressed quantum Hall systems. Subsequently, we demonstrate that the conductivity characteristics of quantum Hall systems can be manipulated using a strong external light.
Our results align with well-established and experimentally
verified theoretical descriptions for undressed systems while providing a more generalized analysis on the conductivity characterstics in quantum Hall systems.
Thus, our model can be applied to accurately interpret the usage of external strong radiation as a tool in nanoscale quantum devices.