### 2021-08-22 paper/v3-release/aps_article.tex
Create release version

% ****** Start of file dressed_quantum_hall.tex ****** %

\documentclass[
 reprint,
 amsmath,amssymb,
 aps,
 prb,
]{revtex4-2}

%===============================================================================
% Import packages
%===============================================================================
% Bold math
\usepackage{bm}
% AMS packages
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
% Physics package
\usepackage{physics}
% SI units package
\usepackage{siunitx}
% Hypertext package
\usepackage[breaklinks=true,colorlinks=true,linkcolor=blue,urlcolor=blue,citecolor=blue]{hyperref}
% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
%===============================================================================
% Bibliography style
\bibliographystyle{apsrev4-2}

\begin{document}

\preprint{APS/123-QED}

% Title of the manuscript
\title{A generalized model for the charge transport properties of
\\dressed quantum Hall systems}

% Authors of the manuscript
\author{Kosala Herath,}
\email[]{kosala.herath@monash.edu}
\author{Malin Premaratne}
\email[]{malin.premaratne@monash.edu}
\affiliation{Advanced Computing and Simulation Laboratory (A$\chi$L), Department of Electrical and Computer Systems Engineering, Monash University, Clayton, Victoria 3800, Australia.}

% Add date
\date{\today}

% Abstract
\begin{abstract}
\input{sections/abstract.tex}
\end{abstract}

% Make title of the manuscript
\maketitle

% Section 01 : Introduction
\section{\label{sec_introduction} Introduction}
\input{sections/1_introduction.tex}

% Section 02 : Schrodinger problem for a dressed quantum Hall system
\section{\label{sec_schrodinger_problem} Schrodinger problem for a dressed quantum Hall system}
\input{sections/2_schrodinger_problem.tex}

% Section 03 : Floquet theory perspectiv
\section{\label{sec_floquet_theory} Floquet theory perspective}
\input{sections/3_floquet_theory.tex}

% Section 04 :
\section{\label{sec_inverse_scattering_time}  Inverse Scattering Time Analysis}
\input{sections/4_inverse_scattering_time.tex}

% Section 05 :
\section{\label{sec_floquet_drude_conductivity} Floquet-Drude Conductivity in a Dressed Quantum Hall System}
\input{sections/5_floquet_drude_conductivity.tex}

% Section 06 :
\section{\label{sec_manipulate_conductivity} Manipulate Conductivity in Quantum Hall Systems}
\input{sections/6_manipulate_conductivity.tex}

% Section 07 :
\section{\label{sec_conclusions} Conclusions}
\input{sections/7_conclusions.tex}

% Acknowledgments
\begin{acknowledgments}
\input{sections/acknowledgments.tex}
\end{acknowledgments}

% Add appendixes
\appendix

% Appendix A : Wave function solutions for a dressed quantum Hall system
\section{\label{appendix_a} Wave function solutions for a dressed quantum Hall system}
\input{sections/appendix_a.tex}

% Appendix B :
\section{\label{appendix_b} Floquet modes and quasienergies}
\input{sections/appendix_b.tex}

% Appendix C :
\section{\label{appendix_c} Floquet-Fermi golden rule for a dressed quantum  Hall system}
\input{sections/appendix_c.tex}

% Appendix D :
\section{\label{appendix_d} Current operators for a dressed quantum Hall system}
\input{sections/appendix_d.tex}

% Bibliography
\bibliography{aps_article}

\end{document}

% ****** End of file dressed_quantum_hall.tex ****** %